---
title: "Introduction to Statistics"
author: "David J. Hunter"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(metathis)
meta() %>%
  meta_description(
    "Resources for my introductory statistics class."
  ) %>%
  meta_name("github-repo" = "djhunter/intro-stats") %>%
  meta_viewport() %>%
  meta_social(
    title = "Introduction to Statistics",
    url = "https://djhunter.github.io/intro-stats/index.html",
    image = "https://ggplot2.tidyverse.org/reference/geom_smooth-1.png",
    image_alt = "Scatterplot",
    og_author = c("David J. Hunter")
  )
```

This site contains slides and other resources that I use in my introductory statistics class.

# Slides 

These slides are meant to accompany the textbook [Introduction to Modern Statistics](https://openintro-ims.netlify.app/), First Edition by Mine Çetinkaya-Rundel and Johanna Hardin, and much of the material contained herein is derivative of this text. As such, these slides are offered with the same Creative Commons ShareAlike license.

1. [Hello Data](slides/01-hellodata.html), 8/29
2. [Association and Experiments](slides/01b-association.html), 8/31
3. [Study Design](slides/02-studydesign.html), 9/2
4. [Getting Started with Data in R](slides/03-intro_data_R.html), 9/5
5. [Exploring Categorical Data](slides/04-catdata.html), 9/7
6. [Exploring Numerical Data](slides/05-numdata.html), 9/9
7. [Exploratory Data Analysis in R](slides/06-explore_data_R.html), 9/12
8. [Review for Test #1](slides/07-test1review.html), 9/14
9. [Linear Regression with a Single Predictor](slides/07a-lin_regression.html), 9/19
10. [Linear Regression with Multiple Predictors](slides/08-multi_regression.html), 9/21
11. [Logistic Regression](slides/09-logit_regression.html), 9/23
12. [Regression Models in R](slides/10-regression_R.html), 9/26
13. [Hypothesis Testing with Randomization](slides/11-hypoth_random.html), 9/28
14. [Confidence Intervals with Bootstrapping](slides/12-bootstrap_ci.html), 9/30
15. [Logistic Regression in R](slides/12b-logit_reg_R.html), 10/3
16. [Review for Test #2](slides/12c-test2review.html), 10/5
17. [Inference with Mathematical Models](slides/13-math_models.html), 10/12
18. [Decision Errors](slides/14-decision_errors.html), 10/14
19. [Simulating Sampling Distributions in R](slides/15-samp_dist_R.html), 10/17
20. [Inference for a Single Proportion](slides/16-infer_1prop.html), 10/19
21. [Inference for Comparing Two Proportions](slides/17-infer_2prop.html), 10/21
22. [Inference for Two-Way Tables](slides/18-infer_chisquare.html), 10/24

# Other Resources

- [Syllabus](syllabus/ma005hunterF22.pdf) (PDF)
- [Textbook Website](https://openintro-ims.netlify.app/)
- [How to install R and RStudio](https://westmont.instructure.com/courses/8503/discussion_topics/44046)

All of the pages and slides on this site were authored in RMarkdown.  The source code is available on GitHub: https://github.com/djhunter/intro-stats

