---
title: "Introduction to Statistics"
author: "David J. Hunter"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(metathis)
meta() %>%
  meta_description(
    "Resources for my introductory statistics class."
  ) %>%
  meta_name("github-repo" = "djhunter/intro-stats") %>%
  meta_viewport() %>%
  meta_social(
    title = "Introduction to Statistics",
    url = "https://djhunter.github.io/intro-stats/index.html",
    image = "https://ggplot2.tidyverse.org/reference/geom_smooth-1.png",
    image_alt = "Scatterplot",
    og_author = c("David J. Hunter")
  )
```

This site contains slides and other resources that I use in my introductory statistics class.

# Slides 

These slides are meant to accompany the textbook [Introduction to Modern Statistics](https://openintro-ims.netlify.app/), First Edition by Mine Çetinkaya-Rundel and Johanna Hardin, and much of the material contained herein is derivative of this text. As such, these slides are offered with the same Creative Commons ShareAlike license.

| Date | Topics |
|:---- |:-------------------------- |
| 8/28 | [Hello Data](slides/01-hellodata.html) |
| 8/30 | [Association and Experiments](slides/01b-association.html) |
|  9/1 | [Study Design](slides/02-studydesign.html) |
| 9/4 | [Getting Started with Data in R](slides/03-intro_data_R.html) |
| 9/6 | [Exploring Categorical Data](slides/04-catdata.html) 
| 9/8 | [Exploring Numerical Data](slides/05-numdata.html) |
| 9/11 | [Exploring Data in R](slides/06b-explore_data_R.html) |
| 9/13 | [Linear Regression with a Single Predictor](slides/07a-lin_regression.html) |

<br>

---
# | 9/15 | [Linear Regression with Multiple Predictors](slides/08-multi_regression.html) |
# | 9/18 | [Linear Regression in R](slides/07b-linear_regression_R.html) 
# | 9/18 | [Multiple Regression in R](slides/10-regression_R.html) |
# | 9/20 |  [Review for Test #1](slides/07c-test1review.html) |
# | 9/22 | Test #1, Chapters 1-8 |
# | 9/25 | [Logistic Regression](slides/09-logit_regression.html) |
# | 9/27 | [Logistic Regression in R](slides/09b-logit_reg_R.html) |
# | 9/29 | [Hypothesis Testing with Randomization](slides/11-hypoth_random.html) |
# | 10/2 |  [Confidence Intervals with Bootstrapping](slides/12-bootstrap_ci.html) |
# | 10/4 | [Inference with Mathematical Models](slides/13-math_models.html) | 
# | 10/6 | [Decision Errors](slides/14-decision_errors.html) |
# | 10/6 | [Simulating Sampling Distributions in R](slides/15-samp_dist_R.html) |
# | 10/11| [Inference for a Single Proportion](slides/16-infer_1prop.html) |
# | 10/13 | [Inference for Comparing Two Proportions](slides/17-infer_2prop.html) | 
# | 10/16 | [Inference for Two-Way Tables](slides/18-infer_chisquare.html) |
# | 10/18 | [Review for Test #2](slides/18c-test2review.html) |
# | 10/20 | Test #2, Chapters 9-18 |
# | 10/23 |  [Inference for a Single Mean](slides/19-infer_onemean.html) |
# | 10/25 | [Bootstrapping and t-tests in R](slides/19b-onemean_R.html) |
# | 10/27 | [Inference for Comparing Two Independent Means](slides/20-infer_twomean.html) |
# | 10/30 | [Inference for Comparing Paired Means](slides/21-infer_pairedmean.html) |
# | 11/1 | [Inference for Comparing Many Means](slides/22-infer_manymeans.html) |
# | 11/3 | [Inference for Numerical Data in R](slides/23b-num_infer_R.html) |
# | 11/6 | [Inference for Linear Regression](slides/24-infer_regression.html) |
# | 11/8 | [Inference for Linear Regression, continued](slides/24-infer_regression.html)  |
# | 11/10 |  [Inference for Multiple Regression](slides/25-infer_multiple.html) |
# | 11/13 | [Inference for Multiple Regression in R](slides/27-mariokart.html) 
# | 11/15 | [Review for Test #3](slides/24c-test3review.html) |
# | 11/17 | Test #3, Chapters 19-25??
# | 11/20 | [Something cool with R?] 
# | 11/27 | [Inference for Logistic Regression](slides/26-infer_logistic.html) |
# | 11/29 | [Inference for Logistic Regression in R](slides/27b-nbarookies.html) |
# Three extra days 
# | 12/8 | [Final Exam Review](slides/27c-test4review.html) |
---

# Other Resources

- [Syllabus](syllabus/ma005hunterF23.pdf) (PDF)
- [Textbook Website](https://openintro-ims.netlify.app/)
- [How to install R and RStudio](slides/installRStudio.html)

All of the pages and slides on this site were authored in RMarkdown.  The source code is available on GitHub: https://github.com/djhunter/intro-stats

