---
title: "Introduction to Statistics"
author: "David J. Hunter"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(metathis)
meta() %>%
  meta_description(
    "Resources for my introductory statistics class."
  ) %>%
  meta_name("github-repo" = "djhunter/intro-stats") %>%
  meta_viewport() %>%
  meta_social(
    title = "Introduction to Statistics",
    url = "https://djhunter.github.io/intro-stats/index.html",
    image = "https://ggplot2.tidyverse.org/reference/geom_smooth-1.png",
    image_alt = "Scatterplot",
    og_author = c("David J. Hunter")
  )
```

This site contains slides and other resources that I use in my introductory statistics class.

# Slides 

These slides are meant to accompany the textbook [Introduction to Modern Statistics](https://openintro-ims.netlify.app/), First Edition by Mine Çetinkaya-Rundel and Johanna Hardin, and much of the material contained herein is derivative of this text. As such, these slides are offered with the same Creative Commons ShareAlike license.

| Date | Topics |
|:---- |:-------------------------- |
| 1/12 | [Hello Data](slides/01-hellodata.html) <br> [Association and Experiments](slides/01b-association.html) |
| 1/19 | [Study Design](slides/02-studydesign.html) <br> [Getting Started with Data in R](slides/03-intro_data_R.html) |
| 1/24 | [Exploring Categorical Data](slides/04-catdata.html) <br> [Exploring Numerical Data](slides/05-numdata.html) |
| 1/26 | [Linear Regression with a Single Predictor](slides/07a-lin_regression.html) |
| 1/31 | [Linear Regression in R](slides/07b-linear_regression_R.html) <br> [Review for Test #1](slides/07c-test1review.html) |
| 2/2 | Test #1, Chapters 1-7 |
| 2/7 | [Linear Regression with Multiple Predictors](slides/08-multi_regression.html) <br> [Multiple Regression in R](slides/10-regression_R.html) |

---
# | 2/9 | [Logistic Regression](slides/09-logit_regression.html) <br> [Logistic Regression in R](slides/12b-logit_reg_R.html) |
#
# 2/14 [Hypothesis Testing with Randomization](slides/11-hypoth_random.html)
#  [Confidence Intervals with Bootstrapping](slides/12-bootstrap_ci.html)
# 
# 2/23 [Inference with Mathematical Models](slides/13-math_models.html)
# [Decision Errors](slides/14-decision_errors.html) 
# 2/28 [Simulating Sampling Distributions in R](slides/15-samp_dist_R.html) 
#  [Inference for a Single Proportion](slides/16-infer_1prop.html) 
#  [Review for Test #2](slides/18b-test3review.html) 
# 3/2 Test #2
# 3/7 [Inference for Comparing Two Proportions](slides/17-infer_2prop.html) 
#  [Inference for Two-Way Tables](slides/18-infer_chisquare.html) 
# 3/9 [Inference for a Single Mean](slides/19-infer_onemean.html)
#  [Inference for Comparing Two Independent Means](slides/20-infer_twomean.html)
# 3/21 [Inference for Comparing Paired Means](slides/21-infer_pairedmean.html)
#  [Inference for Comparing Many Means](slides/22-infer_manymeans.html)
# 3/23 [Inference: Applications](slides/23-applications_infer.html) 
#  [Inference for Numerical Data in R](slides/23b-num_infer_R.html)
# 3/28 [Inference for Linear Regression](slides/24a-infer_regression.html) 
#  [Inference for Linear Regression, continued](slides/24b-infer_regression.html)
# 3/30 [Inference for Multiple Regression](slides/25-infer_multiple.html) 
#  [Inference for Logistic Regression](slides/26-infer_logistic.html)
# 3/30 [Inference for Multiple Regression in R](slides/27-mariokart.html)
#  [Inference for Logistic Regression in R](slides/27b-nbarookies.html)
# 4/4 Data Carpentry
# Review for Test #3
# 4/6 Test #3
# five more content days
# 4/27 [Final Exam Review](slides/27c-test4review.html)
---

# Other Resources

- [Syllabus](syllabus/ma005hunterS23.pdf) (PDF)
- [Textbook Website](https://openintro-ims.netlify.app/)
- [How to install R and RStudio](slides/installRStudio.html)

All of the pages and slides on this site were authored in RMarkdown.  The source code is available on GitHub: https://github.com/djhunter/intro-stats

