---
title: "Introduction to Statistics"
author: "David J. Hunter"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(metathis)
meta() %>%
  meta_description(
    "Resources for my introductory statistics class."
  ) %>%
  meta_name("github-repo" = "djhunter/intro-stats") %>%
  meta_viewport() %>%
  meta_social(
    title = "Introduction to Statistics",
    url = "https://djhunter.github.io/intro-stats/index.html",
    image = "https://ggplot2.tidyverse.org/reference/geom_smooth-1.png",
    image_alt = "Scatterplot",
    og_author = c("David J. Hunter")
  )
```

This site contains slides and other resources that I use in my introductory statistics class.

# Slides 

These slides are meant to accompany the textbook [Introduction to Modern Statistics](https://openintro-ims.netlify.app/), First Edition by Mine Çetinkaya-Rundel and Johanna Hardin, and much of the material contained herein is derivative of this text. As such, these slides are offered with the same Creative Commons ShareAlike license.

| Date | Topics |
|:---- |:-------------------------- |
| 1/9 | [Hello Data](slides/01-hellodata.html) <br> [Association and Experiments](slides/01b-association.html) |
| 1/11 | [Study Design](slides/02-studydesign.html) <br> [Getting Started with Data in R](slides/03-intro_data_R.html) |
| 1/18 | [Exploring Categorical Data](slides/04-catdata.html) <br> [Exploring Numerical Data](slides/05-numdata.html) |
| 1/23 | [Exploring Data in R](slides/06b-explore_data_R.html) <br> [Linear Regression with a Single Predictor](slides/07a-lin_regression.html) |
| 1/25 | [Linear Regression with a Single Predictor (continued)](slides/07a-lin_regression.html#/least-squares-regression) <br> [Linear Regression with Multiple Predictors](slides/08-multi_regression.html) |
| 1/30 | [Regression Models in R](slides/8b-regression_R.html) <br> [Review for Test #1](slides/08c-test1review.html) |
| 2/1 | Test #1, Chapters 1-8 |
| 2/6 | [Logistic Regression](slides/09-logit_regression.html) <br> [Hypothesis Testing with Randomization](slides/11-hypoth_random.html)  |
| 2/8 | [Confidence Intervals with Bootstrapping](slides/12-bootstrap_ci.html) <br> [Bootstrapping in R](slides/12b-bootstrap_ci_R.html)
| 2/13 | [Inference with Mathematical Models](slides/13-math_models.html) <br> [Decision Errors](slides/14-decision_errors.html) |
| 2/15 | [Inference for a Single Proportion](slides/16-infer_1prop.html) |
| 2/22 | [Inference for Comparing Two Proportions](slides/17-infer_2prop.html) <br> [Inference for Two-Way Tables](slides/18-infer_chisquare.html) |
| 2/27 | [Inference on Categorical Variables in R](slides/18a-infer_cat_R.html) <br> [Review for Test #2](slides/18c-test2review.html) |
| 2/29 | Test #2, Chapters 9-18 |
| 3/5 |  [Inference for a Single Mean](slides/19-infer_onemean.html) <br> [Bootstrapping and t-tests in R](slides/19b-onemean_R.html) |
| 3/7 | [Inference for Comparing Two Independent Means](slides/20-infer_twomean.html) |
| 3/19 | [Inference for Comparing Paired Means](slides/21-infer_pairedmean.html) |
| 3/21 | [Inference for Comparing Many Means](slides/22-infer_manymeans.html) |
| 3/26 | [Inference for Numerical Data in R](slides/23b-num_infer_R.html) |
| 3/28 | [Review for Test #3](slides/24c-test3review.html) |
| 4/2 | Test #3, Chapters 16-23 |
| 4/4 | [Inference for Linear Regression](slides/24-infer_regression.html) |

---
# | 4/11 | [Inference for Linear Regression, continued](slides/24-infer_regression.html)  |
# | 4/16 |  [Inference for Multiple Regression](slides/25-infer_multiple.html) |
# | 4/18 | [Inference for Logistic Regression](slides/26-infer_logistic.html) |
# | 4/23 | [Inference for Multiple Regression in R](slides/27-mariokart.html) <br> [Inference for Logistic Regression in R](slides/27b-nbarookies.html) |
# | 4/25 | [Final Exam Review](slides/27c-test4review.html) |
---

# Other Resources

- [Syllabus](syllabus/ma005hunterS24.pdf) (PDF)
- [Textbook Website](https://openintro-ims.netlify.app/)
- [How to install R and RStudio](slides/installRStudio.html)

All of the pages and slides on this site were authored in RMarkdown.  The source code is available on GitHub: https://github.com/djhunter/intro-stats

