---
title: "Chapter 1: Hello Data"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
library(openintro)
library(tidyverse)
library(tools)
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
```

# Overview

- *1.1 Case study: Using stents to prevent strokes*
- *1.2 Data basics*
    - *1.2.1 Observations, variables, and data matrices*
    - *1.2.2 Types of variables*
    - 1.2.3 Relationships between variables
    - 1.2.4 Explanatory and response variables
    - 1.2.5 Observational studies and experiments

# 1.1 Case study: Using stents to prevent strokes

> Does the use of stents reduce the risk of stroke?

-   **Treatment group**. Patients in the treatment group received a stent and medical management. The medical management included medications, management of risk factors, and help in lifestyle modification.
-   **Control group**. Patients in the control group received the same medical management as the treatment group, but they did not receive stents.

Researchers *randomly assigned* 224 patients to the treatment group and 227 to the control group.

## Group activity 

```{r stentStudyResultsDFsummary, echo=FALSE}
stent30_renamed <- stent30 %>% rename(`30 days` = outcome)
stent365_renamed <- stent365 %>% rename(`365 days` = outcome)
stent <- stent30_renamed %>%
  select(-group) %>%
  bind_cols(stent365_renamed) %>% 
  relocate(group) %>%
  mutate(
    group        = fct_rev(group),
    `30 days`    = fct_rev(`30 days`),
    `365 days`   = fct_rev(`365 days`),
  )

stent %>%
  mutate(group = str_to_title(group)) %>%
  pivot_longer(cols = c(`30 days`, `365 days`), 
               names_to = "stage", 
               values_to = "outcome") %>%
  count(group, stage, outcome) %>%
  pivot_wider(names_from = c(stage, outcome), values_from = n) %>%
  kable()
```

1. Of the 224 patients in the treatment group, 45 had a stroke by the end of the first year.
Using these two numbers, compute the proportion of patients in the treatment group who had a stroke by the end of their first year.

2. Compute the proportion of patients in the control group who had a stroke by the end of the first year.

3. What do these two *summary statistics* tell us about the *research question* "Does the use of stents reduce the risk of stroke?"

## Statistical investigations

Steps in a statistical investigation:

- Formulate a *research question*.
- Collect *data*.
- Summarize the data using *summary statistics*.
- *Interpret* the results in the context of the research question.

> Statistics is the study of how best to collect, analyze, and draw conclusions from data.

# 1.2 Data basics

## 1.2.1 Observations, variables, and data frames

- A *data frame* is a table (i.e., a spreadsheet).
    - Each row corresponds to an *observational unit*: the individual entities under consideration.
    - Each column corresponds to a *variable*: properties that can be observed on each observational unit.
    
```{r loan50-df, echo=FALSE}
loan50 %>% 
  select(loan_amount, interest_rate, term, grade, state, total_income, homeownership) %>%
  slice_head(n = 6) %>%
  kable()
```

## `county` Data Frame 

- Observational units: United States counties
- Variables: `r names(county)`

```{r echo=FALSE}
county[1:100,] %>% 
  kbl() %>%
  kable_styling(font_size = 12) %>%
  row_spec(0, background = "darkblue") %>%
  scroll_box(width = "100%", height = "300px")
```

##

```{r county-variables, echo=FALSE}
tribble(
  ~variable,           ~description,
  "name",              "Name of county.",
  "state",             "Name of state.",
  "pop2000",           "Population in 2000.",
  "pop2010",           "Population in 2010.",
  "pop2017",           "Population in 2017.",
  "pop_change",        "Population change from 2010 to 2017 (in percent).",
  "poverty",           "Percent of population in poverty in 2017.",
  "homeownership",     "Homeownership rate, 2006-2010.",
  "multi_unit",        "Multi-unit rate: percent of housing units that are in multi-unit structures, 2006-2010.",
  "unemployment_rate", "Unemployment rate in 2017.",
  "metro",             "Whether the county contains a metropolitan area, taking one of the values yes or no.",
  "median_edu",        "Median education level (2013-2017), taking one of the values below_hs, hs_diploma, some_college, or bachelors.", 
  "per_capita_income", "Per capita (per person) income (2013-2017).",
  "median_hh_income",  "Median household income.",
  "smoking_ban",       "Describes the type of county-level smoking ban in place in 2010, taking one of the values none, partial, or comprehensive."
) %>%
  kbl(row.names = TRUE) %>%
  kable_styling(font_size = 16) %>%
  scroll_box()
```

## 1.2.2 Types of variables

```{r, echo=FALSE}
par_og <- par(no.readonly = TRUE) # save original par
par(mar = rep(0, 4))
plot(c(-0.15, 1.3), 0:1, type = "n", axes = FALSE)

text(0.6, 0.9, "all variables", cex=2)
rect(0.4, 0.8, 0.8, 1)

text(0.25, 0.5, "numerical", cex=2)
rect(0.1, 0.4, 0.4, 0.6)
arrows(0.45, 0.78, 0.34, 0.62, length = 0.08)

text(0.9, 0.5, "categorical", cex=2)
rect(0.73, 0.4, 1.07, 0.6)
arrows(0.76, 0.78, 0.85, 0.62, length = 0.08)

text(0, 0.1, "discrete", cex=2)
rect(-0.17, 0, 0.17, 0.2)
arrows(0.13, 0.38, 0.05, 0.22, length = 0.08)

text(0.39, 0.1, "continuous", cex=2)
rect(0.25, 0, 0.53, 0.2)
arrows(0.35, 0.38, 0.4, 0.22, length = 0.08)

text(0.77, 0.105, "ordinal", cex=2)
rect(0.64, 0, 0.9, 0.2)
arrows(0.82, 0.38, 0.77, 0.22, length = 0.08)

text(1.12, 0.1, "nominal", cex=2)
rect(0.99, 0, 1.25, 0.2)
arrows(1.02, 0.38, 1.1, 0.22, length = 0.08)
par(par_og) # restore original par
```

## Group Activity

1. Data were collected about students in a statistics course. Three variables were recorded for each student: number of siblings, student height, and whether the student had previously taken a statistics course. Classify each of the variables as continuous numerical, discrete numerical, or categorical.

2. An experiment is evaluating the effectiveness of a new drug in treating migraines. A `group` variable is used to indicate the experiment group for each patient: treatment or control. The `num_migraines` variable represents the number of migraines the patient experienced during a 3-month period. Classify each variable as either numerical or categorical?

