---
title: "Chapter 3: Getting Started with Data in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
```

# Overview

- What is R?
- Online Interactive Tutorials
- RStudio

# The R Programming Language

- Free, open-source
- Powerful, professional, widely-used
- Optimized for statistics, ML
- Huge community of support (CRAN)

# Interactive Tutorials 

## Lesson 1: Language of Data

Everyone with a screen (laptop, tablet, or phone), especially the **Operator**, can follow along. Open the first tutorial: 

https://openintro.shinyapps.io/ims-01-data-01/

## Group Activity

The **Recorder** will have some things to write down along the way.

1. How many observations and variables are there in the `email50` data set?
2. What is the name of the function that creates new variables from old ones?
3. Based on the scatterplot of `exclaim_mess` by `num_char` grouped by `spam`, what relationships between these variables do you observe?

## Load Packages

```{r}
library(tidyverse)
library(openintro)
```

## Glimpse

```{r}
glimpse(hsb2)
```

## Count (and pipes)

```{r}
hsb2 %>% 
  count(schtyp)
```

## Filter rows

```{r}
hsb2_public <- hsb2 %>%
  filter(schtyp == "public")
glimpse(hsb2_public)
```

## Count filtered data frame

```{r}
hsb2_public %>%
  count(schtyp)
```

## Making new variables with `mutate`

```{r}
avg_read <- mean(hsb2$read)
hsb2 <- hsb2 %>%
  mutate(read_cat = if_else(read < avg_read,
                            "below average",
                            "at or above average")
         )
hsb2
```

## Make scatterplots

```{r}
ggplot(data = hsb2, aes(x = science, y = math)) + 
  geom_point()
```

## Group by color

```{r}
ggplot(data = hsb2, aes(x = science, y = math, color = prog)) +
  geom_point()
```


# RStudio

Look for an announcement later today on Canvas.

