---
title: "Chapter 3: Getting Started with Data in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
```

# Overview

- What is R?
- Data Carpentry

# The R Programming Language

- Free, open-source
- Powerful, professional, widely-used
- Optimized for statistics, ML
- Huge community of support (CRAN)
- [Installation instructions](installRStudio.html) (hopefully you did this already)

# Data Carpentry

## Work in Pairs

- Please pair up (your choice).
    - At least one person in each pair needs to have a device that can use RStudio (installed or using posit.cloud). This person is the *Operator*.
    - Ideally, everyone has an RStudio-enabled device. If you don't, you can still take notes by keeping track of the code and trying it on your own later.
    
## Data Carpentry

- Fundamental tools for working with data.
    - Open
    - Reproducible
- Live Coding
    - I type, you type.
    - Challenges

## Group Activity

Follow along with the code. Do the challenges. Record answers to any questions asked.

## Load Packages

```{r}
library(openintro)
library(tidyverse)
```

## Glimpse, view, and help

```{r eval=FALSE}
glimpse(hsb2)
view(hsb2)
?hsb2
```

## Count (and pipes)

```{r}
hsb2 %>% 
  count(schtyp)
```

## Challenge

Challenge: How many different values of the `race` variable are there in the data, and how many observations are in each level?

## Make scatterplots

```{r}
hsb2 %>%
  ggplot(aes(x = science, y = math)) +
  geom_point()
```

## Group by color

```{r}
hsb2 %>%
  ggplot(aes(x = science, y = math, color = prog)) +
  geom_point()
```

## Challenge

**Challenge:** Take a look at the email50 data frame. Count the number of emails 
that are spam and the number that are not spam. Make a scatterplot of two of 
the quantitative variables in the data set, colored by one of the categorical
variables.



