---
title: "Chapter 5: Exploring Numerical Data"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
```

# Overview

Chapter 5: Exploring numerical data

- 5.1 Scatterplots for paired data
- 5.2 Dot plots and the mean
- 5.3 Histograms and shape
- 5.4 Variance and standard deviation
- 5.5 Box plots, quartiles, and the median
- 5.6 Robust statistics
- 5.7 Transforming data
- 5.8 Mapping data

# 5.1 Scatterplots for paired data

## Loan Data

```{r, message=FALSE}
library(tidyverse)
library(openintro)
glimpse(loan50)
```

## Scatterplot of amount vs. income

```{r, fig.width=6, fig.height=3.8, fig.align='center'}
loan50 %>%
  ggplot(aes(x = total_income, y = loan_amount)) +
  geom_point()
```

## County Data

```{r, message=FALSE}
glimpse(county)
```

## Scatterplot of income vs poverty rate

```{r, fig.width=6, fig.height=3.5, fig.align='center', message = FALSE, warning=FALSE}
county %>%
  ggplot(aes(x = poverty, y = median_hh_income)) +
  geom_point(alpha = 0.2) +
  geom_smooth()
```



# 5.2 Dot plots and the mean

## One Quantitative Variable

A **dot plot** is a quick way to understand the distribution of a single quantitative variable.

```{r}
loan50$interest_rate
```

## Dot Plot of interest rate

```{r, fig.width=6, fig.height=3.8, fig.align='center', message = FALSE, warning=FALSE}
loan50 %>%
  ggplot(aes(x = interest_rate)) +
  geom_dotplot()
```

## Sample Mean

The **sample mean** $\bar{x}$ is the sum of the values divided by the number of observations.

- The sample mean is a *statistic*.

$$
\bar{x} = \frac{x_1 + x_2 + \cdots + x_n}{n}
$$

```{r}
mean(loan50$interest_rate)
```

## Population Mean

- The `loan50` data set is a *sample* from a much larger *population* of loans.
- The **population mean** is the average of all values in the population.
    - The population mean is a *parameter*.
    - Parameters are generally unknown, but we estimate them with statistics.
        - Our best guess *point estimate* for the population mean interest rate is the sample mean (`r mean(loan50$interest_rate)`).
    - We use the symbol $\mu$ (the Greek letter mu) for the population mean. 
    
## Group Exercise

1. In the dot plot below, pretend that the dots are all physical objects (e.g., weights), and that the horizontal axis is a plank of wood. Where on the plank of wood would you place a support (i.e., fulcrum) so that the whole structure will balance? Everyone should make a guess, then *record the mean* of the guesses in your group.

```{r, fig.width=9, fig.height=3.4, fig.align='center', message = FALSE, warning=FALSE, echo = FALSE}
loan50 %>%
  ggplot(aes(x = interest_rate)) +
  geom_dotplot() + 
  scale_x_continuous(breaks = (1:30))
```



# 5.3 Histograms and shape
# 5.4 Variance and standard deviation
# 5.5 Box plots, quartiles, and the median
# 5.6 Robust statistics
# 5.7 Transforming data
# 5.8 Mapping data
