---
title: "Chapter 6: Exploratory Data Analysis in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(qrencoder)
library(tidyverse)
library(openintro)
```

# Overview


- Graphical representation of two categorical variables
    - Side-by-side bar charts
    - Stacked bar charts
    - To normalize or not to normalize
- Tabular representation of two categorical variables
    - Computation of margins
    - Counts vs proportions
    - Law of total probability
- Graphical representation of one categorical variable
    - Marginal vs conditional
    - Bar chart

# Interactive Tutorials 

## Lesson 1: Visualizing categorical data

Everyone with a screen (laptop, tablet, or phone), especially the **Operator**, can follow along. Open the first tutorial: 

https://openintro.shinyapps.io/ims-02-explore-01/

```{r, echo = FALSE, fig.height=4}
image(qrencode_raster("https://openintro.shinyapps.io/ims-02-explore-01/"), 
      asp=1, col=c("white", "black"), axes=FALSE, 
      xlab="", ylab="")
```

## Group Activity

The **Recorder** should record the answers to the multiple choice questions.
