---
title: "Chapter 6: Exploring Data in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
```

# Data Carpentry

## Work in Pairs

- Please pair up (your choice).
    - At least one person in each pair needs to have a device that can use RStudio (installed or using posit.cloud). This person is the *Operator*.
    - Ideally, everyone has an RStudio-enabled device. If you don't, you can still take notes by keeping track of the code and trying it on your own later.
    
## Group Activity

Follow along with the code. Do the challenges. Record answers to any questions asked.

```{r eval=FALSE}
library(tidyverse)
library(openintro)
data(smoking)
view(smoking)
?smoking

## Categorical Variables

smoking %>% count(marital_status)
smoking %>% count(smoke)
smoking %>% count(marital_status, smoke)
smoking %>%
  ggplot(aes(x = marital_status, fill = smoke)) +
  geom_bar(position = "stack")

## Challenge: Which marital category has the highest proportion of smokers?
## Hints: Try "fill" and/or "dodge" for position

## Challenge: Find another pair of associated categorical variables in this
## data set and illustrate the association.


# Age?
summary(smoking$age)
ggplot(smoking, aes(x = age)) +
  geom_boxplot()
## Challenge: Find another geometry that makes sense for this plot.

## Is there an association between age and whether someone smokes?
ggplot(smoking, aes(x = age, color = smoke)) +
  geom_boxplot()

## Now restrict to just the smokers.
smokers <- smoking %>%
  filter(smoke == "Yes")
glimpse(smoking)
glimpse(smokers)
ggplot(smokers, aes(x = amt_weekends)) +
  geom_boxplot()

## Challenge: Q. Do men and women smokers differ in how much they smoke on 
## weekends? Make a plot and record (log sheet) a
## conclusion sentence.

## plot a single numerical variable
smoking %>%
  ggplot(aes(x = age)) +
  geom_boxplot()
smoking %>%
  ggplot(aes(x = age)) +
  geom_histogram()
smoking %>%
  ggplot(aes(x = age)) +
  geom_density()

## Challenge: Is there an association between age
## and whether someone smokes? (record a sentence) 
## Try adding smoke as an aesthetic for the box, hist, and
## density plots. (e.g., fill = smoke)

```
