---
title: "Chapter 7: Linear Regression with a Single Predictor"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(qrencoder)
library(tidyverse)
library(openintro)
```

# Overview

- 7 Linear regression with a single predictor
    - 7.1 Fitting a line, residuals, and correlation
        - 7.1.1 Fitting a line to data
        - 7.1.2 Using linear regression to predict possum head lengths
        - 7.1.3 Residuals
        - 7.1.4 Describing linear relationships with correlation
    - 7.2 Least squares regression
        - 7.2.1 Gift aid for freshman at Elmhurst College
        - 7.2.2 An objective measure for finding the best line
        - 7.2.3 Finding and interpreting the least squares line
        - 7.2.4 Extrapolation is treacherous
        - 7.2.5 Describing the strength of a fit
        - 7.2.6 Categorical predictors with two levels
    - 7.3 Outliers in linear regression

# 7.1 Fitting a line, residuals, and correlation

## 7.1.1 Fitting a line to data

## 7.1.2 Using linear regression to predict possum head lengths

## 7.1.3 Residuals

## 7.1.4 Describing linear relationships with correlation

# 7.2 Least squares regression

## 7.2.1 Gift aid for freshman at Elmhurst College

## 7.2.2 An objective measure for finding the best line

## 7.2.3 Finding and interpreting the least squares line

## 7.2.4 Extrapolation is treacherous

## 7.2.5 Describing the strength of a fit

## 7.2.6 Categorical predictors with two levels

# 7.3 Outliers in linear regression

## Comments on Homework Problems

https://openintro-ims.netlify.app/

```{r, echo = FALSE, fig.height=4}
image(qrencode_raster("https://openintro-ims.netlify.app/"), 
      asp=1, col=c("white", "black"), axes=FALSE, 
      xlab="", ylab="")
```

