---
title: "Linear Regression Models in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
library(openintro)
theme_set(theme_bw())
```

# Pairs Today

- Please pair up (your choice).
    - At least one person in each pair needs to have a device that can use RStudio (installed or using rstudio.cloud). This person is the *Operator*.
    - Ideally, everyone has an RStudio-enabled device. If you don't, you can still take notes by keeping track of the code and trying it on your own later.
    
# Data Carpentry

- Fundamental tools for working with data.
    - Open
    - Reproducible
- Live Coding
    - I type, you type.
    - Challenges

## Group Exercises

Do all the challenges.

```{r, eval = FALSE}
library(tidyverse)
library(openintro)

# Look at the duke_forest data frame
duke_forest
glimpse(duke_forest)
view(duke_forest)

# Make a new data frame: 
#   Get rid of address, hoa, url columns

homes <- duke_forest %>%
  select(-address, -hoa, -url)
homes
view(homes)

# predicting price

## Scatterplot of price vs. area

ggplot(homes, aes(x = area, y = price)) +
  geom_point()

## Problem: We have some very influential observations
# with areas above 4500. Let's remove them

homes_small_area <- homes %>%
  filter(area < 4500)

ggplot(homes_small_area, aes(x = area, y = price)) + 
  geom_point()

## Challenge: Use homes:
# 1. Make a scatterplot of price vs lot.
# 2. Filter out lot sizes above 1.0. Call the new data frame homes_small_lot.
# 3. Make a scatterplot of price vs lot in your new data frame.
# 4. Which do you think has a stronger correlation (lot or area)?

## Which is the better predictor?

model1 <- lm(price ~ area, data = homes)
summary(model1)

model1s <- lm(price ~ area, data = homes_small_area)
summary(model1s)

# Write down the regression equation, interpret regression slope.

## Challenge: make a model predicting price from lot. (model2 and model2s, 
# using homes and homes_small_lot, respectively. ) Based on the R^2 values,
# which predictor (area or lot) is more strongly associated with house price?

## Add the regression lines to the plots
ggplot(homes, aes(x = area, y = price)) +
  geom_point() +
  geom_smooth(method = "lm")

## Challenge: 
### 1. Add regression line to price vs lot scatterplot.
### 2. Modify 1 aesthetics to color by the cooling variable.
### solution

## use the models to make a prediction
new_house <- data.frame(area = 1600, lot = 0.25)
predict(model1, new_house)
predict(model1s, new_house)
predict(model2, new_house)
predict(model2s, new_house)

## Categorical response variable (OPTIONAL)

#   Add a variable anypark: 0 if no spaces, 1 otherwise

homes$parking
unique(homes$parking)

homespark <- homes %>%
  mutate(anypark = factor(ifelse(parking == "0 spaces", 0, 1))) 
view(homespark)

## Predict anypark from lot size

ggplot(homespark, aes(x = lot, y = anypark)) +
  geom_boxplot()
```
