---
title: "Chapter 8: Linear Regression with Multiple Predictors"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
library(openintro)
library(kableExtra)
loans <- loans_full_schema %>%
  mutate(
    credit_util = total_credit_utilized / total_credit_limit,
    bankruptcy  = as.factor(if_else(public_record_bankrupt == 0, 0, 1)),
    verified_income = droplevels(verified_income)
    ) %>%
  rename(credit_checks = inquiries_last_12m) %>%
  select(interest_rate, verified_income, debt_to_income, credit_util, bankruptcy, term, credit_checks, issue_month) 
theme_set(theme_bw())
```

# Overview

- 8 Linear regression with multiple predictors
    - 8.1 Indicator and categorical predictors
    - 8.2 Many predictors in a model
    - 8.3 Adjusted R-squared
    - 8.4 Model selection
        - 8.4.1 Stepwise selection
        - 8.4.2 Other model selection strategies

#  8.1 Indicator and categorical predictors

## Indicator variables

```{r}
glimpse(loans)
```

`bankruptcy` is an *indicator* (categorical 0 or 1) variable with two levels:

- 0 means the borrower has not had a bankruptcy in the past.
- 1 means the borrower has had a bankruptcy in the past.

## Regression with categorical predictor

Does having a past bankruptcy predict the interest rate you get?

```{r}
model1 <- lm(interest_rate ~ bankruptcy, data = loans)
summary(model1)$coefficients
```

Regression equation: $\widehat{\texttt{interest_rate}} = 12.34 + 0.74 \times \texttt{bankruptcy}$

- If `bankruptcy` is 0, $\widehat{\texttt{interest_rate}} = 12.34$.
- If `bankruptcy` is 1, $\widehat{\texttt{interest_rate}} = 12.34 + 0.74 = 13.08$.

## Predictors with more than two levels

`verified_income` has three levels:

- *Verified:* Both the source and amount of income were verified.
- *Source Verified:* Only the source was verified, not the amount.
- *Not Verified:* The income was not verified.

```{r}
model2 <- lm(interest_rate ~ verified_income, data = loans)
summary(model2)$coefficients
```

## Interpreting a 3-level predictor

```{r, echo = FALSE}
model2 <- lm(interest_rate ~ verified_income, data = loans)
summary(model2)$coefficients
```

\[
\begin{aligned}
\widehat{\texttt{interest_rate}} &= 11.10 \\
&+ 1.42 \times \texttt{verified_income}_{\texttt{Source Verified}} \\
&+ 3.25 \times \texttt{verified_income}_{\texttt{Verified}}
\end{aligned}
\]

- The *reference level* is `Not Verified`: $\widehat{\texttt{interest_rate}} = 11.10$
- If `verified_income` is `Source Verified`, you have to add 1.42.
- If `verified_income` is `Verified`, you have to add 3.25.

## Boxplot of interest rates by verification

```{r, fig.width=6, fig.height=3.8, fig.align='center', message = FALSE, warning=FALSE}
ggplot(data = loans, aes(x = interest_rate, y = verified_income)) +
  geom_boxplot()
```

## Group Discussion

\[
\begin{aligned}
\widehat{\texttt{interest_rate}} &= 11.10 \\
&+ 1.42 \times \texttt{verified_income}_{\texttt{Source Verified}} \\
&+ 3.25 \times \texttt{verified_income}_{\texttt{Verified}}
\end{aligned}
\]

1. Do the variables `interest_rate` and `verified_income` appear to be associated? How?

2. Can you think of a possible *confounding variable* that might account for this association?

## R-squared in model #1

```{r}
summary(model1)
```

## R-squared in model #2

```{r}
summary(model2)
```

## How much does the model tell us?

- Model #1 only explains about 0.2% of the variation in interest rates.
- Model #2 only explains about 6% of the variation in interest rates.

#  8.2 Many predictors in a model

## Variable descriptions

```{r, echo = FALSE}
tribble(
  ~variable,         ~description,
  "interest_rate",   "Interest rate on the loan, in an annual percentage.",
  "verified_income", "Categorical variable describing whether the borrower's income source and amount have been verified, with levels `Verified`, `Source Verified`, and `Not Verified`.",
  "debt_to_income",  "Debt-to-income ratio, which is the percentage of total debt of the borrower divided by their total income.",
  "credit_util",     "Of all the credit available to the borrower, what fraction are they utilizing. For example, the credit utilization on a credit card would be the card's balance divided by the card's credit limit.",
  "bankruptcy",      "An indicator variable for whether the borrower has a past bankruptcy in their record. This variable takes a value of `1` if the answer is *yes* and `0` if the answer is *no*.",
  "term",            "The length of the loan, in months.",
  "issue_month",     "The month and year the loan was issued, which for these loans is always during the first quarter of 2018.", 
  "credit_checks",   "Number of credit checks in the last 12 months. For example, when filing an application for a credit card, it is common for the company receiving the application to run a credit check.",
) %>%
  kbl()
```

## Multiple Regression Model

\[
\begin{aligned}
\widehat{\texttt{interest_rate}} &= b_0 \\
&+ b_1 \times \texttt{verified_income}_{\texttt{Source Verified}} \\
&+ b_2 \times \texttt{verified_income}_{\texttt{Verified}} \\
&+ b_3 \times \texttt{debt_to_income} \\
&+ b_4 \times \texttt{credit_util} \\
&+ b_5 \times \texttt{bankruptcy} \\
&+ b_6 \times \texttt{term} \\
&+ b_9 \times \texttt{credit_checks} \\
&+ b_7 \times \texttt{issue_month}_{\texttt{Jan-2018}} \\
&+ b_8 \times \texttt{issue_month}_{\texttt{Mar-2018}}
\end{aligned}
\]

## Multiple Regression in R

```{r}
model3 <- lm(interest_rate ~ verified_income + debt_to_income + credit_util +
               bankruptcy + term + credit_checks + issue_month, data = loans)
summary(model3)$coefficients
```

## Equation for fitted model

\[
\begin{aligned}
\widehat{\texttt{interest_rate}} &= 1.89 \\
&+ 1.00 \times \texttt{verified_income}_{\texttt{Source Verified}} \\
&+ 2.56 \times \texttt{verified_income}_{\texttt{Verified}} \\
&+ 0.02 \times \texttt{debt_to_income} \\
&+ 4.90 \times \texttt{credit_util} \\
&+ 0.39 \times \texttt{bankruptcy} \\
&+ 0.15 \times \texttt{term} \\
&+ 0.23 \times \texttt{credit_checks} \\
&+ 0.05 \times \texttt{issue_month}_{\texttt{Jan-2018}} \\
&- 0.04 \times \texttt{issue_month}_{\texttt{Mar-2018}}
\end{aligned}
\]

## Group Exercise

<div class = "column-left">
\[
\scriptsize{
\begin{aligned} 
&\widehat{\texttt{interest_rate}} = 1.89 \\
&+ 1.00 \times \texttt{verified_income}_{\texttt{Source Verified}} \\
&+ 2.56 \times \texttt{verified_income}_{\texttt{Verified}} \\
&+ 0.02 \times \texttt{debt_to_income} \\
&+ 4.90 \times \texttt{credit_util} \\
&+ 0.39 \times \texttt{bankruptcy} \\
&+ 0.15 \times \texttt{term} \\
&+ 0.23 \times \texttt{credit_checks} \\
&+ 0.05 \times \texttt{issue_month}_{\texttt{Jan-2018}} \\
&- 0.04 \times \texttt{issue_month}_{\texttt{Mar-2018}}
\end{aligned} }
\]
</div>

<div class = "column-right">
1. Interpret the coefficients of `credit_checks` and `bankruptcy`.

2. Predict the interest rate of a borrower with the following characteristics:

- Income is Source Verified
- Debt-to-income is 21
- Credit utilized is 0.32
- Never has had a bankruptcy
- Term of loan is 36 months
- Number of credit checks is 3
- Loan is issued in March

</div>

#  8.3 Adjusted R-squared

## Full summary of regression model

```{r}
summary(model3)
```

##  Multiple and Adjusted R-squared

- *Multiple* $R^2$ estimates the percentage of variability in the response variable that is explained by the model.
    - This combination of variables explains about 26% of the variability in interest rate.
- *Adjusted* $R^2$ takes into account the complexity of the model.
    - Adding explanatory variables will always increase the multiple $R^2$, but you might be *overfitting* the data (using too complex a model).
    - If you add an explanatory variable to a model and the Adjusted $R^2$ goes *down*, you probably have too many variables.

#  8.4 Model selection


##  8.4.1 Stepwise selection

The `step` function in R attempts to eliminate explanatory variables that aren't needed.

- It computes a score (the AIC) that measures how good the model is. (lower is better)
- It tries eliminating variables that *decrease* the AIC. 

## Using `step`

```{r}
step(model3)
```

## Improved model

```{r}
model4 <- lm(formula = interest_rate ~ verified_income + debt_to_income + 
    credit_util + bankruptcy + term + credit_checks, data = loans)
summary(model4)
```

## Too many variables?

- Model #3 (with all the variables) has an adjusted $R^2$ of 0.2597.
- Model #4 (without issue_month) has an adjusted $R^2$ of 0.2598.
    - This suggests that the variable `issue_month` isn't helping the predictive power of the model, so we shouldn't include it.


