---
title: "Chapter 11: Hypothesis Testing with Randomization"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
library(openintro)
library(kableExtra)
theme_set(theme_bw())
```

# Overview

- 11 Hypothesis testing with randomization
    - 11.1 Sex discrimination case study
        - 11.1.1 Observed data
        - 11.1.2 Variability of the statistic
        - 11.1.3 Observed statistic vs. null statistics
    - 11.2 Opportunity cost case study
        - 11.2.1 Observed data
        - 11.2.2 Variability of the statistic
        - 11.2.3 Observed statistic vs. null statistics
    - 11.3 Hypothesis testing
        - 11.3.1 The US court system
        - 11.3.2 p-value and statistical significance


# 11.1 Sex discrimination case study

**Research Question:** Are female employees discriminated against in promotion decisions made by male managers?

- 48 male supervisors were given a simulated personnel file
- Response variable: should the employee be promoted?
- Explanatory variable: sex of employee (randomly assigned)

## 11.1.1 Observed data

```{r}
glimpse(sex_discrimination)
```

```{r}
table(sex_discrimination)
```

## Group Discussion

```{r}
table(sex_discrimination)
```

```{r}
prop.table(table(sex_discrimination), margin = 1)
```

1. Describe the association between the variables `sex` and `decision`.

2. Do you think these proportions indicate evidence of discrimination, or could they just have occurred by random chance? How are you deciding?

## Statistics, Parameters, Hypotheses

Statistics:

- $\hat{p}_F \approx 0.58333$ is the proportion of females that were promoted.
- $\hat{p}_M = 0.875$ is the proportion of males that were promoted.
- $\hat{p}_M - \hat{p}_F \approx 0.292$ is the difference in proportions.

. . . 

Parameters: 

- $\pi_F$ is the actual probability that a female will be promoted.
- $\pi_M$ is the actual probability that a male will be promoted.
- $\pi_M - \pi_F$ is the difference in proportions.

. . . 

Hypotheses:

- $H_0: \pi_M - \pi_F = 0$ (*Null Hypotheses:* no discrimination)
- $H_A: \pi_M - \pi_F > 0$ (*Alternative Hypothesis:* discrimination against women)

## 11.1.2 Variability of the statistic

- Observed Statistic: $\hat{p}_M - \hat{p}_F \approx 0.292$
- $H_0: \pi_M - \pi_F = 0$ (*Null Hypotheses:* no discrimination)
- $H_A: \pi_M - \pi_F > 0$ (*Alternative Hypothesis:* discrimination against women)

The observed statistic is in the direction of the alternative hypothesis, but there are two possibilities:

- $H_A$ is true, we have evidence of discrimination.
- $H_0$ is true, and we are just observing random variation.

## Modeling Random Variation

```{r}
summary(sex_discrimination)
```

- Make 35 "promoted" cards (red) and 13 "not promoted" cards (blue)
- Randomly assign 24 cards male and 24 female
- Compute the difference in proportions several times.
- Compare to the observed difference in proportions.

## {data-background-image="https://openintro-ims.netlify.app/images/sex-rand-03-shuffle-1-sort.png" data-background-size="contain"}

## 11.1.3 Observed statistic vs. null statistics

- The *observed statistic* is 0.292.
- We repeat the card shuffling 100 times and compute 100 *null statistics*.
    - The null statistic simulates what would happen if $H_0$ were true.
- If we rarely see null statistics as extreme as the observed statistic, we have evidence that $H_0$ is false.

## {data-background-image="https://openintro-ims.netlify.app/11-foundations-randomization_files/figure-html/sex-rand-dot-plot-1.png" data-background-size="contain"}

## p-value

The simulation-based **p-value** is the proportion of simulated statistics that were as extreme or more extreme than the observed statistic.

- We simulated 100 statistics.
- Only 2 of them were more extreme that 0.292.
- p-value $= 2/100 = 0.02$

The smaller the p-value, the more evidence you have against $H_0$ in favor of $H_A$.

## {data-background-image="" data-background-size="contain"}

## Dolphin study

https://www.rossmanchance.com/applets/2021/chisqshuffle/ChiSqShuffle.htm?dolphins=1


# 11.2 Opportunity cost case study

## 11.2.1 Observed data

## 11.2.2 Variability of the statistic

## 11.2.3 Observed statistic vs. null statistics

# 11.3 Hypothesis testing

## 11.3.1 The US court system 

## 11.3.2 p-value and statistical significance
