---
title: "Chapter 12: Confidence Intervals with Bootstrapping"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
library(openintro)
library(kableExtra)
library(qrencoder)
theme_set(theme_bw())
```

# Overview

- 12 Confidence intervals with bootstrapping
    - 12.1 Medical consultant case study
        - 12.1.1 Observed data
        - 12.1.2 Variability of the statistic
    - 12.2 Tappers and listeners case study
        - 12.2.1 Observed data
        - 12.2.2 Variability of the statistic
    - 12.3 Confidence intervals
        - 12.3.1 Plausible range of values for the population parameter
        - 12.3.2 Bootstrap confidence interval


# 12 Confidence intervals with bootstrapping

- Use for *estimating the value of an unknown parameter*.
    - Get a range (i.e., interval) of plausible values.
- Best used when data have been generated through *random sampling*.
    - The range corresponds to the variability of the data in a systematic way.
    - As we repeatedly sample from a population, what range of sample statistics should we expect to see?

# 12.1 Medical consultant case study

**Research Question:** Do medical consultants lead patients to better outcomes?

## 12.1.1 Observed data

- The average complication rate for liver donor surgeries is about 10%.
- The clients of a particular consultant had only 3 complications in 62 surgeries.
    - $\hat{p} = 3/62 \approx 0.048$
    
## 12.1.2 Variability of the statistic

- We want to know the actual probability of a complication for the clients of this consultant:
    - parameter: $\pi$
- Our best estimate of $\pi$ is the *statistic* $\hat{p} \approx 0.048$.
    - But this is just one *sample*. How would our estimate differ if we took another sample?
- **Bootstrapping:** We can approximate the variability in different samples by taking samples from a population that looks like an "infinite" version of the original sample data.

## {data-background-image="https://openintro-ims.netlify.app/images/boot1prop1.png" data-background-size="contain"}

## {data-background-image="https://openintro-ims.netlify.app/images/boot1prop2.png" data-background-size="contain"}

## {data-background-image="https://openintro-ims.netlify.app/images/boot1prop3.png" data-background-size="contain"}

## Sampling with replacement

- In practice, we sample from the original data *with replacement*.
- This means that when we select an observation, we add it to the "resample" and then put it back in the original data (we replace it).
- We *resample* multiple times and obtain many different resampled proportions.
- A dot plot of these *bootstrapped* proportions approximates a distribution that shows how the statistic varies.

## {data-background-image="https://openintro-ims.netlify.app/images/boot1prop4.png" data-background-size="contain"}

## {data-background-image="https://openintro-ims.netlify.app/images/boot1propboth.png" data-background-size="contain"}

## {data-background-image="https://openintro-ims.netlify.app/12-foundations-bootstrapping_files/figure-html/MedConsBSSim-1.png" data-background-size="contain"}

## Bootstrap 95% confidence interval

- 95% of the bootstrapped proportions are between 0 and 0.113.
- If you repeated this study many times, 95% of the observed proportions would fall between 0 and 0.113.
    - Interpretation: We are "95% confident" that the actual proportion is between 0 and 0.113.
    - Based on the consultant's data, it is *plausible* that the actual proportion of complications with her patients is 0.10. This number is within the range of "typical" values: from 0 to 0.113.
    
# Reeses Pieces

How many orange Reeses Pieces are in a typical bag of 40?

## Observed Data

- Consider a bag of 40 Reeses Pieces as a random sample from the population of all Reeses Pieces.
- You obtain a bag of 40 Reeses Pieces and discover that 20 are orange. ($\hat{p} = 0.5$).

## Variability of the statistic

- Your bag of 40 Reeses Pieces had 20 that were orange. ($\hat{p} = 0.5$)
- Consider an infinite candy machine containing Reeses Pieces, 50% of which are orange 
- We can construct a *bootstrap sample* by drawing 40 Reeses Pieces out of this machine.
- Each bootstrap sample has a $\hat{p}_\text{boot}$. Keep track of these in a dotplot to see how the sample proportion varies.

## Reeses Pieces Applet

https://www.rossmanchance.com/applets/2021/oneprop/OneProp.htm?candy=1

```{r, echo = FALSE, fig.height=4}
image(qrencode_raster("https://www.rossmanchance.com/applets/2021/oneprop/OneProp.htm?candy=1"), 
      asp=1, col=c("white", "black"), axes=FALSE, 
      xlab="", ylab="")
```

## Group Activity

1. Set the *Probability of orange* to 0.5, the *Number of candies* to 40, the *Number of samples* to 1, and choose the *Proportion of orange* statistic. 

2. Click *Draw Samples*. How many orange did you get? What is $\hat{p}_\text{boot}$? Does it equal $\hat{p}$? Is it close?

3. Now change *Number of samples* to 99 and click *Draw Samples* again. When it finishes, make sure *Total Samples* is 100.

4. The dotplot that appears should have 100 dots. Use this dotplot to estimate a bootstrap 90% confidence interval for the proportion of orange Reeses Pieces.

5. Now repeat steps 1-4, but suppose that your bag of candy had 400 Reeses Pieces, 200 of which were orange. How did the confidence interval change? 

# 12.2 Tappers and listeners case study

## 12.2.1 Observed data

In a sample of 120 listeners, only 3 were able to correctly guess the tune that was being tapped. ($\hat{p} = 0.025$)

## 12.2.2 Variability of the statistic

Simulation:

- Get a bag of 120 marbles, 3 of which are red, and the rest white.
- Draw from the bag 120 times, *replacing* the marble each time, and keep track of the number of red.
- Compute $\hat{p}_\text{boot}$
- Repeat 10000 times

$$
0.0417 \quad 0.025 \quad 0.025 \quad 0.0083 \quad 0.05 \quad 0.0333 \quad 0.025 \quad 0 \quad 0.0083 \quad 0
$$

## {data-background-image="https://openintro-ims.netlify.app/12-foundations-bootstrapping_files/figure-html/tappers-bs-sim-1.png" data-background-size="contain"}

# 12.3 Confidence intervals

## 12.3.1 Plausible range of values for the population parameter

- Given a sample of data, we can construct a *confidence interval* for $\pi$.
    - The center of this interval will be at $\hat{p}$.
- This confidence interval is a range of *plausible values* for the population parameter.
    - Based on our data, it is *plausible* that the unknown parameter lies somewhere in our confidence interval.


## 12.3.2 Bootstrap confidence interval

>- The 95% bootstrap confidence interval for the parameter $\hat{p}$ can be obtained directly using the ordered $\hat{p}_\text{boot}$ values.

>- Consider the sorted $\hat{p}_\text{boot}$ values. Call the 2.5% bootstrapped proportion value "lower," and call the 97.5% bootstrapped proportion value "upper."

>- The 95% confidence interval is given by: (lower, upper)

>- We can use different *confidence levels* other than 95%. The larger the confidence level, the wider the interval.

>- Larger samples give narrower confidence intervals, all else being equal.