---
title: "Bootstrapping in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
theme_set(theme_bw())
```

# Pairs Today

- Please pair up (your choice).
    - At least one person in each pair needs to have a device that can use RStudio (installed or using rstudio.cloud). This person is the *Operator*.
    - Ideally, everyone has an RStudio-enabled device. If you don't, you can still take notes by keeping track of the code and trying it on your own later.
    
# Data Carpentry

## Group Exercises

Do all the challenges.

```{r, eval = FALSE}
library(tidyverse)

gss <- read_csv("http://math.westmont.edu/data/gss.csv")

glimpse(gss)

gss %>% 
  count(church)

ggplot(gss, aes(x = church)) +
  geom_bar()

gss %>%
  summarize(phat = sum(church == "often")/300)

## Create a random resample
samp1 <- gss %>%
  slice_sample(n = 300, replace = TRUE)

samp1 %>%
  count(church)
  
# Challenge: 
# 1. Repeat the above two commands (make a sample, then count) a 
#   few times. Does the sample change?
# 2. Summarize to compute p_hat for this sample

## Make lots of p_hats

makephat <- function(i) {
  gss %>%
  slice_sample(n = 300, replace = TRUE) %>%
  summarize(p_hat = sum(church == "often")/300)
}

## Try it out:
makephat()
makephat()

## Use map to repeat the function:
map(1:5, makephat)

map(1:5, makephat) %>%
  bind_rows()

phatDF <- map(1:100, makephat) %>%
  bind_rows()

ggplot(phatDF, aes(x = p_hat)) +
  geom_histogram()

## Challenge: 
## 1. Modify the above code to create a sampling distribution with 1000 p_hats.
## 2. Try using geom_density instead of geom_histogram.

## Create a confidence interval

phatDF %>%
  summarize(p025 = quantile(p_hat, 0.025),
            p975 = quantile(p_hat, 0.975))

## Interpretation: If we repeated the survey many times, 95% of the 
## phats would fall between these two numbers.

gss %>%
  count(church)

## Simulated data: sample size of 3000 instead of 300.
biggss <- data.frame(
  church = c(rep("often", 1360), rep("rarely", 1640))
)

### Challenge: Redo the bootstrapping simulation, but this time
#              use the biggss simulated data. How does the 95% CI change?

```

