---
title: "Review for Test #2"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(qrencoder)
library(tidyverse)
library(openintro)
```

# General Advice 

See the [study guide](https://westmont.instructure.com/courses/8503/files/602415) on Canvas.

- Review the slides and your class notes. 
    - As you review, follow along with the *textbook*.
    - As you review, keep a list of key terms and concepts.
    - Make sure you can answer all the *group activity questions*.
- Review your homework, keeping in mind that these have been graded very leniently and not all of your answers will be correct, even if you didn't have points deducted.
- Do some of the *odd-numbered problems* in the textbook.
    - The answers to the odd-numbered questions are in the back of the book.
- Study the *Chapter Review* sections at the end of each chapter.

# Review Problems

## Comments on Homework Problems

https://openintro-ims.netlify.app/

```{r, echo = FALSE, fig.height=4}
image(qrencode_raster("https://openintro-ims.netlify.app/"), 
      asp=1, col=c("white", "black"), axes=FALSE, 
      xlab="", ylab="")
```

## Group Activity

- Chapter 7, Exercise 21
- Chapter 8, Exercises 7, 9
- Chapter 9, Exercise 5
- Chapter 11, Exercise 7
- Chapter 12, Exercises 3, 7

Recorder and Presenter may have some assigned tasks.

