---
title: "Inference on Categorical Variables in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
theme_set(theme_bw())
```

# Pairs Today

- Please pair up (your choice).
    - At least one person in each pair needs to have a device that can use RStudio (installed or using posit.cloud). This person is the *Operator*.
    - Ideally, everyone has an RStudio-enabled device. If you don't, you can still take notes by keeping track of the code and trying it on your own later.
    
# Review

## Two-proportion test

- Explanatory variable: Categorical, 2 levels
- Response variable: Categorical, 2 levels
- Observed statistic: $\hat{p}_1 - \hat{p}_2$
- Null Hypothesis: $H_0: p_1 - p_2 = 0$
- Alternative Hypothesis: $H_A: p_1 - p_2 \neq 0$
    - Small P-values give you evidence in favor of $H_A$ (groups are different)
- Can also compute confidence intervals for $p_1 - p_2$.

## Chi-squared test

- Explanatory variable: Categorical, $\geq$ 2 levels
- Response variable: Categorical, $\geq$ 2 levels
- Observed statistic: $\chi^2$
- Null Hypothesis: No association (independent)
- Alternative Hypothesis: Some association (not independent)
    - Small P-values give you evidence of an association (groups are not all the same)
- Confidence intervals don't apply.

# Data Carpentry

## General Social Survey Data

| Variable name | Value | Description |
| ------------ | ----------------------- | ---------------------------- |
| `party` | `Dem`, `Rep`, or `other` | Subject's political party preference. (Democrat, Republican, or Independent/Other) |
| `belief` | `athiest`, `theist`, or `unsure` | Subject's belief in God. |
| `church ` | `often` or `rarely` | How frequently the subject attends church or other religious services. |
| `opera` | `like` or `not` | Does the subject like opera music? (Like or do not like) |
| `metal` | `like` or `not` | Does the subject like heavy metal music? |
| `happy` | `yes` or `no`  | Does the subject identify as generally happy? |
| `educ`  | Number of years of school | Subject's education level |
| `age` | Subject's age in years | |
| `words` | Number of words correct | Subject's score on a 10 point vocabulary quiz. |
| `status` | SES score on a 100 point scale | Subject's socioeconomic status |
| `income` | Subject's income in dollars | |

## Group Exercises

Do all the challenges.

```{r, eval = FALSE}
library(tidyverse)

gss <- read_csv("http://math.westmont.edu/data/gss.csv")

gss %>%
  ggplot(aes(x = metal)) +
    geom_bar()

## Challenge: incorporate opera into the above plot. Do you think
## the plot suggests an association between opera and metal? Record
## a sentence.

# make a table
gss %>%
  select(metal, opera) %>%
  table()

## Challenge: Make two new variables:
## p1hat = proportion of metal-likers who like opera
## p2hat = proportion of metal-dislikers who like opera
## Then compute p1hat - p2hat

gss %>%
  select(metal, opera) %>%
  table() %>% 
  prop.test()

gss %>%
  select(metal, opera) %>%
  table() %>% 
  chisq.test()

## Challenge: Is political party associated with church attendance? 
## Make a bar plot: Record what you observe.
## Do a chisq.test. Record what you conclude.
## Try a prop.test. What happens. Why?

## Challenge: Find two categorical variables in the data
## set that have a significant association. Make a bar plot
## and do an appropriate test. Record the variables you used.

# Challenge: Are churchgoers happier? Answer with a bar plot and an 
# appropriate test/CI.
```

