---
title: "Bootstrapping in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
theme_set(theme_bw())
```

# Pairs Today

- Please pair up (your choice).
    - At least one person in each pair needs to have a device that can use RStudio (installed or using rstudio.cloud). This person is the *Operator*.
    - Ideally, everyone has an RStudio-enabled device. If you don't, you can still take notes by keeping track of the code and trying it on your own later.
    
# Review

# Data Carpentry

## General Social Survey Data

| Variable name | Value | Description |
| ------------ | ----------------------- | ---------------------------- |
| `party` | `Dem`, `Rep`, or `other` | Subject's political party preference. (Democrat, Republican, or Independent/Other) |
| `belief` | `athiest`, `theist`, or `unsure` | Subject's belief in God. |
| `church ` | `often` or `rarely` | How frequently the subject attends church or other religious services. |
| `opera` | `like` or `not` | Does the subject like opera music? (Like or do not like) |
| `metal` | `like` or `not` | Does the subject like heavy metal music? |
| `happy` | `yes` or `no`  | Does the subject identify as generally happy? |
| `educ`  | Number of years of school | Subject's education level |
| `age` | Subject's age in years | |
| `words` | Number of words correct | Subject's score on a 10 point vocabulary quiz. |
| `status` | SES score on a 100 point scale | Subject's socioeconomic status |
| `income` | Subject's income in dollars | |

## Group Exercises

Do all the challenges.

```{r, eval = FALSE}
```

