---
title: "Chapter 20: Inference for comparing two independent means"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
library(openintro)
library(kableExtra)
theme_set(theme_bw())
set.seed(1423)
```

# Overview

- 20.1 Randomization test for the difference in means
    - 20.1.1 Observed data
    - 20.1.2 Variability of the statistic
    - 20.1.3 Observed statistic vs. null statistics
- 20.2 Bootstrap confidence interval for the difference in means
    - 20.2.1 Observed data
    - 20.2.2 Variability of the statistic
- 20.3 Mathematical model for testing the difference in means
    - 20.3.1 Observed data
    - 20.3.2 Variability of the statistic
    - 20.3.3 Observed statistic vs. null statistics
- 20.4 Mathematical model for estimating the difference in means
    - 20.4.1 Observed data
    - 20.4.2 Variability of the statistic

# 20.1 Randomization test for the difference in means

- Null Hypothesis: No association
    - $H_0: \mu_1 - \mu_2 = 0$
- Alternative Hypothesis: Some association
    - $H_A: \mu_1 - \mu_2 \neq 0$

## 20.1.1 Observed data

- Suppose you split up a class randomly into two groups: One group takes exam form A, the other takes exam form B. 
- The group taking exam A averaged slightly better than the exam B group: 1.45 points better, on average.

>- Observed Statistic: $\bar{x}_A - \bar{x}_B = 1.45$
>- $H_0: \mu_A - \mu_B = 0$
>- $H_A: \mu_A - \mu_B \neq 0$


## {data-background-image="https://openintro-ims.netlify.app/20-inference-two-means_files/figure-html/boxplotTwoVersionsOfExams-1.png" data-background-size="contain"}

## 20.1.2 Variability of the statistic

## {data-background-image="https://openintro-ims.netlify.app/images/rand2means.png" data-background-size="contain"}

## {data-background-image="https://openintro-ims.netlify.app/20-inference-two-means_files/figure-html/randexams-1.png" data-background-size="contain"}

## 20.1.3 Observed statistic vs. null statistics

## {data-background-image="https://openintro-ims.netlify.app/20-inference-two-means_files/figure-html/randexamspval-1.png" data-background-size="contain"}

# Sleep Deprivation {data-background="https://d2ck0sxsjau14o.cloudfront.net/wp-content/uploads/2017/08/sleepy-man.jpg"}

## Sleep Deprivation

Researchers Stickgold, James, and Hobson investigated delayed effects of sleep deprivation on learning in a study published in Nature Neuroscience (2000). The goal of the study was to see whether the improvement scores tend to be higher for the unrestricted sleep treatment than for the sleep deprivation treatment.

## Statistics and Hypotheses

Unrestricted-sleep group's improvement scores (milliseconds): 

* 25.20, 14.50, -7.00, 12.60, 34.50, 45.60, 11.60, 18.60, 12.10, 30.50

Sleep-deprived group's improvement scores (milliseconds): 

* -10.70, 4.50, 2.20, 21.30, -14.70, -10.70, 9.60, 2.40, 21.80, 7.20, 10.00

\[ 
\bar{x}_\text{unrestricted} - \bar{x}_\text{deprived} = 15.92 \\[0.2in]
\begin{align}
H_0: \mu_\text{unrestricted} - \mu_\text{deprived} &= 0 \\
H_a: \mu_\text{unrestricted} - \mu_\text{deprived} &\neq 0
\end{align}
\]

## Simulation with Cards

- Unrestricted-sleep: 25.20, 14.50, -7.00, 12.60, 34.50, 45.60, 11.60, 18.60, 12.10, 30.50
- Sleep-deprived:  -10.70, 4.50, 2.20, 21.30, -14.70, -10.70, 9.60, 2.40, 21.80, 7.20, 10.00

Because the null hypothesis asserts that improvement score is not associated with sleep condition, we will assume that the 21 subjects would have had exactly the same improvement scores as they did, regardless of which sleep condition group (unrestricted or deprived) the subject had been assigned.

- 21 cards; numbers on cards
- Deal into two piles: 10 and 11

## Group Activity

1. Open the [Multiple Means applet](http://www.rossmanchance.com/applets/AnovaShuffle.htm?hideExtras=2). Check the *Show Shuffle Options* box, select the *Plot* display, and press *Shuffle Responses.*  **Describe what happens** when you press *Shuffle Responses*. How is the location of the next point on the dotplot determined?

2. Use a total of at least 1000 repetitions, and **record a p-value.**  **Record a sentence** summarizing the strength of evidence in the context of this study.

3. **Record** an SE **95% confidence interval** for $\mu_\text{unrestricted} - \mu_\text{deprived}$. Is zero a plausible value? 

# 20.2 Bootstrap confidence interval for the difference in means

## {data-background-image="https://openintro-ims.netlify.app/images/bootmeans2means.png" data-background-size="contain"}

## 20.2.2 Variability of the statistic

## {data-background-image="https://openintro-ims.netlify.app/20-inference-two-means_files/figure-html/bootexamsci-1.png" data-background-size="contain"}

# 20.3 Mathematical model for testing the difference in means

The *Central Limit Theorem* applies to differences in means as well. Conditions:

- Independent observations
- Large samples without extreme outliers, or smaller samples with symmetric populations.

## 20.3.1 Observed data

Data on birth weight and mother smoking habit:

```{r}
births14 %>%
  group_by(habit) %>%
  summarize(xbar = mean(weight),
            s = sd(weight),
            n = n())
```

## {data-background-image="https://openintro-ims.netlify.app/20-inference-two-means_files/figure-html/babySmokePlotOfTwoGroupsToExamineSkew-1.png" data-background-size="contain"}

## Hypotheses

- $H_0 :$ There is no difference in average birth weight for newborns from mothers who did and did not smoke. 
    - $\mu_n - \mu_s = 0$
- $H_A:$ There is some difference in average newborn weights from mothers who did and did not smoke. 
    - $\mu_n - \mu_s \neq 0$
    
## 20.3.3 Observed statistic vs. null statistics

T-statistic for independent means:

$$
T = \frac{(\bar{x}_1 - \bar{x}_2) - 0}{\sqrt{s_1^2/n_1 + s_2^2/n_2}}
$$
Degrees of freedom equals minimum of $n_1-1$ and $n_2-1$.

## Group Exercise

```{r echo=FALSE}
births14 %>%
  group_by(habit) %>%
  summarize(xbar = round(mean(weight), 2),
            s = round(sd(weight), 2),
            n = n()) %>%
  drop_na() 
```

1. Compute a T-score for this hypothesis test. 
$$
T = \frac{(\bar{x}_1 - \bar{x}_2) - 0}{\sqrt{s_1^2/n_1 + s_2^2/n_2}}
$$


2. Convert the T-score to a p-value using `pt`. Degrees of freedom equals minimum of $n_1-1$ and $n_2-1$.

3. Write a sentence interpreting your conclusion in the context of the study.

# Using the mathematical model in R

## Glimpse the data

```{r}
glimpse(births14)
```

## Formulas in R

- Response ~ Explanatory
- `weight ~ habit`

## Two-sample t-test in R

```{r}
t.test(weight ~ habit, data = births14)
```

## Confidence interval interpretation

> We are 95% confident that the mean weight of babies born to nonsmokers is between 0.3 and 0.9 pounds more than the mean weight of babies born to smokers.

## Some bonus Data Carpentry

Revisit the GSS data set:

```{r, eval=FALSE}
library(tidyverse)
gss <- read_csv("http://math.westmont.edu/data/gss.csv")
```

P-hack: Find a numerical response and a categorical explanatory (2 levels) with a significant association. Make a side-by-side boxplot and do a t.test.

