---
title: "Chapter 23: Applications of Inference"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
library(openintro)
library(kableExtra)
theme_set(theme_bw())
set.seed(1423)
```

# Overview

- 23.1 Recap: Computational methods
    - 23.1.1 Randomization
    - 23.1.2 Bootstrapping
- 23.2 Recap: Mathematical models
    - 23.2.1 z-procedures
    - 23.2.2 t-procedures
- 23.3 Case study: Redundant adjectives
    - 23.3.1 Exploratory analysis
    - 23.3.2 Confidence interval for a single mean
    - 23.3.3 Paired mean test
    - 23.3.4 Two independent means test
    
# 23.1 Recap: Computational methods

- Two types of inference for *an unknown parameter*.
    - Hypothesis Tests 
        - Assume $H_0$ about your parameter, and assess how extreme your observed statistic would be.
        - Extreme statistics (low P-values) indicate evidence against $H_0$ in favor of $H_A$.
    - Confidence intervals
        - Measure the observed statistic, and compute a margin of error.
        - The confidence interval represents a range of plausible values for the unknown parameter.

## 23.1.1 Randomization (Hypothesis Tests)

- Write appropriate hypotheses.
- Compute the observed statistic of interest.
- Permute the data repeatedly, each time, recalculating the statistic of interest.
- Compute the proportion of times the permuted statistics are as extreme as or more extreme than the observed statistic, this is the p-value.
- Make a conclusion based on the p-value, and write the conclusion in context and in plain language so anyone can understand the result.

## 23.1.2 Bootstrapping (Confidence Intervals)

- Repeatedly resample the original data, with replacement, using the same sample size as the original data.
- For each resample, calculate the statistic of interest.
- Calculate the confidence interval using one of the following methods:
    - Bootstrap percentile interval: Obtain the endpoints representing the middle (e.g., 95%) of the bootstrapped statistics. 
    - Bootstrap standard error (SE) interval: Find the SE of the bootstrapped statistics. The confidence interval will be the observed statistic plus or minus some multiple (e.g., 2) of SEs.
- Put the conclusions in context and in plain language so even non-statisticians can understand the results.

# 23.2 Recap: Mathematical models

| Explanatory | Response | Model |
| ----------- | -------- | ----- |
| Categorical ($\geq$ 2 levels) | Categorical ($\geq$ 2 levels) | Chi-squared |
| none | Numeric | One-sample t-test |
| Categorical (2 levels) | Numeric | Two-sample t-test |
| Categorical (2 levels, paired) | Numeric | Paired t-test |
| Categorical (> 2 levels) | Numeric | ANOVA F-test |

## Practice Problems: Identify the Scenario

Complete the quiz at the following Google Form:

https://forms.gle/ibCCkVJrt6EubG4b9


# 23.3 Case study: Redundant adjectives

- What factors affect whether people use "redundant adjectives"?
    - Context
    - Native Language
- Experiment: show two images to 22 native English speakers (undergraduates from University College London) and 22 native Spanish speakers (undergraduates from the Universidad de las Islas Baleares). 

## {data-background-image="https://openintro-ims.netlify.app/images/redundant-adjectives-blue-triangle.png" data-background-size="contain"}


## 23.3.1 Exploratory analysis

```{r include=FALSE}
# analysis based on 
# https://osf.io/fqnms/ > Production data > AnalyzeProduction.R
# one row per question
redundant_individual <- read_csv("data/ENGLISH-SPANISH PRODUCTION.csv") %>%
  mutate(
    items = if_else(Question > 10, 16, 4),
    items = as.factor(items),
    adjective = if_else(`Color Response` == 1, "redundant", "not redundant")
    ) %>%
  select(-`Color Response`)
# one row per individual
redundant <- redundant_individual %>%
  group_by(Language, Subject, items) %>%
  summarise(
    n_questions = n(),
    redundant_perc = sum(adjective == "redundant") * 100 / n_questions,
    .groups = "drop"
    )
```

```{r}
glimpse(redundant)
```

## Redundant adjectives: boxplots

```{r, fig.height=4, fig.width=6, fig.align='center'}
ggplot(redundant, aes(x = items, y = redundant_perc, color = Language)) + geom_boxplot()
``` 


## Redundant adjectives: violin plots

```{r, fig.height=4, fig.width=6, fig.align='center'}
ggplot(redundant, aes(x = items, y = redundant_perc, color = Language)) + geom_violin()
``` 

## 23.3.2 Confidence interval for a single mean

Mean redundant adjective usage: English speakers, 4 items

```{r}
redundant %>%
  filter(Language == "English",
         items == 4) %>%
  t.test(redundant_perc ~ 1, data = .)
```

## {data-background-image="https://openintro-ims.netlify.app/23-inference-applications_files/figure-html/boot-eng-4-viz-1.png" data-background-size="contain"}

# Group Activty

For each research question, (1) determine the appropriate *scenario*, (2) state *hypotheses* in symbols, and (3) describe the *observed statistic* in words.

- Do the data provide convincing evidence of a difference in mean redundant adjective usage percentages between sparse (4 item) and dense (16 item) displays for English speakers?

- How does redundant adjective usage differ between English speakers and Spanish speakers?

## 23.3.3 Paired mean test

```{r}
redundant %>%
  filter(Language == "English") %>%
  pivot_wider(names_from = items, 
              values_from = redundant_perc,
              names_prefix = "item") %>%
  {t.test(.$item16, .$item4, paired = TRUE)}
```

## {data-background-image="https://openintro-ims.netlify.app/23-inference-applications_files/figure-html/eng-viz-1.png" data-background-size="contain"}

## Item comparison (English speakers)

```{r, fig.height=4, fig.width=6, fig.align='center'}
redundant %>%
  filter(Language == "English") %>%
  ggplot(aes(x = items, y = redundant_perc)) +
    geom_boxplot()
```

## Item comparison (English speakers)

```{r, fig.height=4, fig.width=6, fig.align='center'}
redundant %>%
  filter(Language == "English") %>%
  ggplot(aes(x = items, y = redundant_perc)) +
    geom_violin()
```


## 23.3.4 Two independent means test


## Two mean t-test (4 items)

```{r}
redundant %>%
  filter(items == 4) %>%
  t.test(redundant_perc ~ Language, data = .)
```



## Two mean t-test (16 items)

```{r}
redundant %>%
  filter(items == 16) %>%
  t.test(redundant_perc ~ Language, data = .)
```


## {data-background-image="https://openintro-ims.netlify.app/23-inference-applications_files/figure-html/compare-lang-viz-1.png" data-background-size="contain"}

## English vs. Spanish (4 items)

```{r, fig.height=4, fig.width=6, fig.align='center'}
redundant %>%
  filter(items == 4) %>%
  ggplot(aes(x = redundant_perc, fill = Language)) +
    geom_histogram(binwidth = 10, position = "dodge")
```


## English vs. Spanish (16 items)

```{r, fig.height=4, fig.width=6, fig.align='center'}
redundant %>%
  filter(items == 16) %>%
  ggplot(aes(x = redundant_perc, fill = Language)) +
    geom_histogram(binwidth = 10, position = "dodge")
```
