---
title: "Inference for Numerical Data in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
library(openintro)
theme_set(theme_bw())
```

# Pairs Today

- Please pair up (your choice).
    - At least one person in each pair needs to have a device that can use RStudio (installed or using rstudio.cloud). This person is the *Operator*.
    - Ideally, everyone has an RStudio-enabled device. If you don't, you can still take notes by keeping track of the code and trying it on your own later.
    
# Review

## Categorical Explanatory, Numerical Response

- Two-sample $t$-test
    - Explanatory variable has 2 levels.
    - Data is *not* paired; groups are *independent*.
    - Statistic: $\bar{x}_1 - \bar{x}_2$
- ANOVA $F$-test
    - Explanatory variable has more than 2 levels.
    - Statistic: $F = \frac{MSG}{MSE}$
    - Omnibus test
        - If significant, follow up with Tukey HSD.
        - If not significant, there is no association. Don't follow up.
    
# Data Carpentry

## Group Exercises

Do all the challenges.

```{r, eval = FALSE}
library(tidyverse)
library(openintro)

data(smoking)
glimpse(smoking)
?smoking
summary(smoking)

summary(smoking$age)
```

---

```{r, eval = FALSE}
ggplot(smoking, aes(x = age)) +  geom_boxplot() 
ggplot(smoking, aes(x = age)) +  geom_histogram() 
ggplot(smoking, aes(x = age)) +  geom_density() 

ggplot(smoking, aes(x = age, fill = smoke)) +  geom_boxplot() 
ggplot(smoking, aes(x = age, fill = smoke)) +  geom_histogram(position = "dodge") 
ggplot(smoking, aes(x = age, fill = smoke)) +  geom_density(alpha = 0.3) 

t.test(age ~ smoke, data = smoking)
```

---

```{r, eval = FALSE}
## Let's restrict our attention to just people who smoke.

smokers <- smoking %>%
  filter(smoke == "Yes")

summary(smokers)

ggplot(smokers, aes(x = amt_weekends)) + geom_boxplot()

## Challenge: Do men and women differ in how much they smoke on weekends? ## Make a plot and do a t-test.
```

---

```{r, eval = FALSE}
## Multiple means: Grouping variables with more than 2 levels

summary(smokers$marital_status)

ggplot(smokers, aes(x = amt_weekends, fill = marital_status)) + geom_boxplot()
mstat.aov <- aov(amt_weekends ~ marital_status, data = smokers)
anova(mstat.aov)

## Challenge: Can you find a factor in the data set that is
## associated with how much someone smokes on weekends? 
## Try some. Make a boxplot, and then do an anova. 
## Is there evidence of an association? If so, follow up with TukeyHSD.
```

