---
title: "Review for Test #4"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
library(openintro)
```

# General Advice 

- See the [study guide](https://westmont.instructure.com/courses/8503/files/625761). The test covers chapters 19-23, but lots of the material depends on Chapters 11-18, so you might want to go back and do some selected review of those chapters.
- Review the slides and your class notes. 
    - As you review, follow along with the *textbook*.
    - As you review, keep a list of key terms and concepts.
    - Make sure you can answer all the *group activity questions*.
- Review your homework, keeping in mind that these have been graded very leniently.
- Do some of the *odd-numbered problems* in the textbook.
    - The answers to the odd-numbered questions are in the back of the book.
- Study the *Chapter Review* sections at the end of each chapter.

## Equations provided

- Generic SE confidence interval
    - $\mbox{point estimate} \pm 2 \cdot SE$
- One-sample t-test
    - $\displaystyle{T = \frac{\bar{x} - \text{null value}}{s/\sqrt{n}}}$, $df = n-1$ 
- One mean t-interval
    - $\displaystyle{\bar{x}  \pm  t^{\star}_{df} \times \frac{s}{\sqrt{n}}}$ 
- Two-sample t-test (independent means)
    - $\displaystyle{T = \frac{(\bar{x}_1 - \bar{x}_2) - 0}{\sqrt{s_1^2/n_1 + s_2^2/n_2}}}$, $df = \text{min}(n_1-1, n_2-1)$ 
    
## Equations provided (continued)

- Two-sample t-interval
    - $\displaystyle{\bar{x}_1 - \bar{x}_2 \pm t^\star_{df} \times \sqrt{\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}}}$ 
- Paired t-test
    - $\displaystyle{T = \frac{\bar{x}_\text{diff} - 0}{s_\text{diff}/\sqrt{n}}}$, $df = n-1$ 
- Paired t-interval
    - $\displaystyle{\bar{x}_\text{diff}  \pm  t^{\star}_{df} \times \frac{s_\text{diff}}{\sqrt{n}}}$ 
- ANOVA F-test
    - $\displaystyle{F = \frac{MSG}{MSE}}$, $df_1 = k-1$, $df_2 = n-k$

# Review Problems

## Questions on Homework Problems?

https://openintro-ims.netlify.app/

## Group Activity

- Chapter 19, Exercises 13, 17
- Chapter 20, Exercises 5, 6
- Chapter 21, Exercises 4, 7 
- Chapter 22, Exercises 5 (see next slides)

Recorder and Presenter may have some assigned tasks.

## Chapter 22, Exercise 5

```{r}
chick.aov <- aov(weight ~ feed, data = chickwts)
anova(chick.aov)
```

## Follow-up test

```{r}
TukeyHSD(chick.aov)
```



