---
title: "Inference for Multiple Regression in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
library(openintro)
theme_set(theme_bw())
```

# Pairs Today

- Please pair up (your choice).
    - At least one person in each pair needs to have a device that can use RStudio (installed or using rstudio.cloud). This person is the *Operator*.
    - Ideally, everyone has an RStudio-enabled device. If you don't, you can still take notes by keeping track of the code and trying it on your own later.
    
# Case Study: Mario Kart {data-background="https://images.nintendolife.com/screenshots/9506/900x.jpg" data-background-size="contain"}

## Mario Kart Data

>- Observational Units: 141 Ebay auctions of Mario Kart Wii.
>- Response (Outcome) Variable: total price of the auction.
>- Several possible explanatory variables: duration of auction in days, condition (new/used), number of bids, starting price, seller rating, etc.
>- Can we determine how total price is related to these variables, controlling for other variables? For example:
>    - All other characteristics held constant, are longer auctions associated with higher or lower prices?
>    - On average, how much more do buyers tend to pay for additional Wii wheels (plastic steering wheels that attach to the Wii controller) in auctions?

# Data Carpentry

## Group Exercises

Do all the challenges.

```{r, eval = FALSE}
library(tidyverse)
library(openintro)
library(caret)
```

---

