---
title: "Inference for Multiple Regression in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
library(openintro)
theme_set(theme_bw())
```

# Pairs Today

- Please pair up (your choice).
    - At least one person in each pair needs to have a device that can use RStudio (installed or using rstudio.cloud). This person is the *Operator*.
    - Ideally, everyone has an RStudio-enabled device. If you don't, you can still take notes by keeping track of the code and trying it on your own later.
    
# Case Study: Mario Kart {data-background="https://images.nintendolife.com/screenshots/9506/900x.jpg" data-background-size="contain"}

## Mario Kart Data

>- Observational Units: 143 Ebay auctions of Mario Kart Wii.
>- Response (Outcome) Variable: total price of the auction.
>- Several possible explanatory variables: duration of auction in days, condition (new/used), number of bids, starting price, seller rating, etc.
>- Can we determine how *total price* is related to these variables, controlling for other variables? For example:
>    - All other characteristics held constant, are longer auctions associated with higher or lower prices?
>    - On average, how much more do buyers tend to pay for additional Wii wheels (plastic steering wheels that attach to the Wii controller) in auctions?

# Data Carpentry

## Group Exercises

Do all the challenges.

```{r, eval = FALSE}
library(tidyverse)
library(openintro)
library(caret)
```

---

```{r, eval = FALSE}
## View the data

view(mariokart)

?mariokart
```

---

```{r, eval = FALSE}
## Check for outliers

ggplot(mariokart, aes(x = total_pr)) + geom_histogram()

## What is the deal with those?

mariokart %>%
  filter(total_pr > 100) %>% 
  view()

## Remove outliers

mk <- mariokart %>%
  filter(total_pr < 100)

ggplot(mk, aes(x = total_pr)) + geom_histogram()
```

---

```{r, eval = FALSE}
## Linear regression model
ggplot(mk, aes(x = total_pr, y = cond)) + geom_boxplot()
ggplot(mk, aes(x = total_pr, y = cond)) + geom_violin()

mmod1 <- lm(total_pr ~ cond, data = mk)
summary(mmod1)
```

---

```{r, eval = FALSE}
## Control for Wii wheels
glimpse(mk)

ggplot(mk, aes(x = wheels, y = total_pr)) + geom_point()
ggplot(mk, aes(x = wheels, y = total_pr, color = cond)) + geom_point()

## Challenge: Control for wheels by adding it as a predictor. Call your new
## model mmod2. Compare with mmod1. What is the price difference between
## new and used versions?
```

---

```{r, eval = FALSE}
## Cross Validation
set.seed(3456)
tc <- trainControl(method = "cv", number = 5)
train(total_pr ~ cond, data = mk, method = "lm", trControl = tc)
train(total_pr ~ cond + wheels, data = mk, method = "lm", trControl = tc)

## Challenge: Find a model for predicting total_pr with a lower RMSE.

## Challenge: For the best model you found, 
## 1. Are all the predictors significant? 
## 2. Record the adjusted R^2.

## Challenge: make some plots illustrating the important predictors that you found.
```
