---
title: "Inference for Logistic Regression in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
library(openintro)
theme_set(theme_bw())
```

# Pairs Today

- Please pair up (your choice).
    - At least one person in each pair needs to have a device that can use RStudio (installed or using rstudio.cloud). This person is the *Operator*.
    - Ideally, everyone has an RStudio-enabled device. If you don't, you can still take notes by keeping track of the code and trying it on your own later.
    
# {data-background="https://s3media.247sports.com/Uploads/Assets/933/733/9733933.png" data-background-size="contain"}

## NBA Rookie Data

>- Observational Units: 2621 NBA Rookies.
>- Response (Outcome) Variable: Are they still in the NBA after 5 years? (TRUE/FALSE)
>- Several possible explanatory variables: games played, minutes per game, points per game, rebounds, assists, etc.
>- Can we predict which types of players will remain in the NBA?
>    - Which predictors are important?
>    - Which predictors are correlated?

# Data Carpentry

## Group Exercises

Do all the challenges.

---

```{r, eval = FALSE}
library(tidyverse)
library(caret)
```

---

```{r, eval = FALSE}
## Load and inspect the data
nba <- read_csv("https://math.westmont.edu/data/nba.csv")
view(nba)

ggplot(nba, aes(x = PTS)) + geom_histogram()
ggplot(nba, aes(x = PTS, y = TARGET_5Yrs)) + geom_point()
ggplot(nba, aes(x = PTS, y = TARGET_5Yrs)) + geom_boxplot()
```

---

```{r, eval = FALSE}
## Logistic regression models
bbmod1 <- glm(TARGET_5Yrs ~ PTS, family = "binomial", data = nba)
summary(bbmod1)

bbmod2 <- glm(TARGET_5Yrs ~ 3P, family = "binomial", data = nba)
## Need backticks:
bbmod2 <- glm(TARGET_5Yrs ~ `3P`, family = "binomial", data = nba)

summary(bbmod2)

# Challenge: Improve bbmod2 by controlling for 3 point attempts (3PA). Do
# the p-values change? Can you explain why?

```

---

```{r, eval = FALSE}
## Correlated predictors?
cor(nba[,7:31])
round(cor(nba[,7:31]), 1)

ggplot(nba, aes(x = `3PA`, y = `3P`)) + geom_point()
ggplot(nba, aes(x = `3P%`)) + geom_histogram()
```

---

```{r, eval = FALSE}
## Cross Validation

tc <- trainControl(method = "cv", number = 10)
bbcv1 <- train(TARGET_5Yrs ~ PTS, 
               data = nba, method = "glm", family = "binomial", trControl = tc)
# Error: Needs to be a factor
nba$TARGET_5Yrs <- factor(nba$TARGET_5Yrs)
bbcv1 <- train(TARGET_5Yrs ~ PTS, 
               data = nba, method = "glm", family = "binomial", trControl = tc)
confusionMatrix(bbcv1)
bbcv2 <- train(TARGET_5Yrs ~ `3P` + `3PA`, 
               data = nba, method = "glm", family = "binomial", trControl = tc)
confusionMatrix(bbcv2)
bbcv2

## Challenge: Find a model with a better prediction accuracy.
## For the best model you find, generate a logistic regression 
## table. Which predictors are most important? 
```

---

```{r, eval = FALSE}
## Search for best model?
nba_complete <- nba %>%
  drop_na() %>%
  select(-Player, -PlayerID, -Pos, -Tm)
fullmod <- glm(TARGET_5Yrs ~ ., family = "binomial", data = nba_complete)
summary(fullmod)
step(fullmod)
```
