---
title: "Summary of Inference Methods in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
theme_set(theme_bw())
```

# Pairs Today

- Please pair up (your choice).
    - At least one person in each pair needs to have a device that can use RStudio (installed or using posit.cloud). This person is the *Operator*.
    - Ideally, everyone has an RStudio-enabled device. If you don't, you can still take notes by keeping track of the code and trying it on your own later.
    
# Review

## Summary of Explanatory/Response Scenarios

# Data Carpentry

## General Social Survey Data

| Variable name | Value | Description |
| ------------ | ----------------------- | ---------------------------- |
| `belief` | `athiest`, `theist`, or `unsure` | Subject's belief in God. |
| `major` | `STEM`, `other`, or `none` | Subject's college major (STEM = Science, Technology, Engineering, or Mathematics) |
| `boyGirl` | `corr` or `incDK` | Subject's answer to the question, "It is the father’s gene that decides whether the baby is a boy or a girl. (Is that true or false?)" (correct or incorrect/don't know) |
| `atoms` | `corr` or `incDK`  | Subject's answer to the question, "Electrons are smaller than atoms. (Is that true or false?)" |
| `earth` | `corr` or `incDK`  | Subject's answer to the question, "How long does it take for the Earth to go around the Sun: one day, one month, or one year?" |
| `church ` | `often` or `rarely` | How frequently the subject attends church or other religious services. |
| `age` | Subject's age in years | |
| `words` | Number of words correct | Subject's score on a 10 point vocabulary quiz. |
| `educ`  | Number of years of school | Subject's education level |
| `status` | SES score on a 100 point scale | Subject's socioeconomic status |
| `income` | Subject's income in dollars | |

## Group Exercises

Do all the challenges.

```{r, eval = FALSE}
library(tidyverse)

gss <- read_csv("http://math.westmont.edu/data/gss2.csv")

gss %>%
  ggplot(aes(x = church, fill = atoms)) +
    geom_bar(position = "fill")

?prop.test

gss %>%
  select(church, atoms) %>%
  table() %>%
  prop.test()

## Interpretation: Churchgoers are significantly less likely to know 
## that electrons are smaller than atoms. 

?chisq.test

## Challenge: Conduct a chi-square test for an appropriate choice of 
## variables in this data set. Record an interpretation.

?t.test

## Challenge: Conduct a two-sample t-test for an appropriate choice of 
## variables and record an interpretation.

?aov
?TukeyHSD

## Challenge: Conduct an ANOVA F-test for an appropriate choice of 
## variables, and follow up appropriately. Record an interpretation.

?lm

## Challenge: Fit a linear regression model for an appropriate choice of 
## variables. Record an interpretation of the p-values.

?glm

## Challenge: Fit a logistic regression model for an appropriate choice of 
## variables. Record an interpretation of the p-values.
```

