---
title: "Regression Models in R"
output: 
  revealjs::revealjs_presentation:
    fig_width: 14 
    fig_height: 7
    self_contained: true
    theme: night
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 540
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
options(width = 100)
library(tidyverse)
library(openintro)
theme_set(theme_bw())
```

# Pairs Today

- Please pair up (your choice).
    - At least one person in each pair needs to have a device that can use RStudio (installed or using rstudio.cloud). This person is the *Operator*.
    - Ideally, everyone has an RStudio-enabled device. If you don't, you can still take notes by keeping track of the code and trying it on your own later.
    
# Data Carpentry

- Fundamental tools for working with data.
    - Open
    - Reproducible
- Live Coding
    - I type, you type.
    - Challenges

## Group Exercises

Do all the challenges.

```{r, eval = FALSE}
library(tidyverse)
library(openintro)

# Look at the duke_forest data frame
duke_forest
glimpse(duke_forest)
view(duke_forest)

# Make a new data frame: 
#   Get rid of address, hoa, url columns

homes <- duke_forest %>%
  select(-address, -hoa, -url)
homes
view(homes)

## Single variable models
homes %>%
  ggplot(aes(x = area, y = price)) +
  geom_point() +
  geom_smooth(method = "lm")

homes %>%
  ggplot(aes(x = lot, y = price)) +
  geom_point() + 
  geom_smooth(method = "lm")

model1 <- lm(price ~ area, data = homes)
summary(model1)

model2 <- lm(price ~ lot, data = homes)
summary(model2)

## Multiple regression model

model3 <- lm(price ~ area + lot, data = homes)
summary(model3)

## use the model to make a prediction
predict(model3, data.frame(area = 1600, lot = 0.25))

## Challenge: multiple regression model
## Make a new model (model4) that predicts price from 
## area, lot, bed, bath, year_built, cooling
## Is this an improvement over model 3?

## eliminate some variables
step(model4)

## Challenge: make a new model (model5) using the step results. 
##    1. Compare adjusted R^2 to model4. Which model is better?
##    2. Predict the price of a house with 2000 sq feet area, 
##       lot size 0.5, 3 baths, built in 1950, with central cooling.

```
