---
title: "Section 10.6"
date: "December 4, 2019"
output: 
  word_document:
  revealjs::revealjs_presentation:
    fig_width: 14
    fig_height: 7
    self_contained: true
    theme: black
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 600
---

# Section 10.6: Logistic Regression and Factorial Anova

## Questions on Chapter 10 Investigation?

## Recall Chapter 8-9 Investigation

```{r, fig.height=4.3, fig.width=6, fig.align='center'}
drivers <- read.table("http://math.westmont.edu/ma5/carstatus.txt", header = TRUE)
plot(Behavior ~ Status, data = drivers)
```

## No evidence?

>- We found little to no evidence for an association between car status and yielding behavior.
>- Categorical $\rightarrow$ Categorical (Chi-square, p-value $\approx$ 0.2)
>- However, we observed a trend, so maybe it makes sense to view car status as *Quantitative* (1,2,3,4,5)
>- **Operator:** Look at https://stats.idre.ucla.edu/other/mult-pkg/whatstat/ and decide which test to use.

# Logistic Regression

## Predicting probabilities 

A *logistic* curve has the following shape:

```{r, echo=FALSE, fig.height=5}
plot(function(x) {exp(x)/(1+exp(x))}, xlim = c(-10,10), ylab = "predicted probability", xlab="quantitative explanatory variable")
```

## Status coded as categorical (factor)

```{r}
drivers
```

## Change to numeric (quantitative)

```{r}
drivers$Status <- as.numeric(drivers$Status)
drivers
```

## Logistic Regression in R

```{r}
d.logit <- glm(Behavior ~ Status, data = drivers, family = binomial)
summary(d.logit)
```

## Significance and predictions

- The logistic regression model had a p-value of 0.03, indicating strong evidence of an association between driving behavior and yielding behavior.
- The prediction equation has the form 
$$
\displaystyle \hat{P}(x) = \frac{e^{b_0 + b_1x}}{1 + e^{b_0 + b_1x}}
$$ 
where $x$ is a status value and $b_0$ and $b_1$ are the logistic regression coefficients.

## Logistic prediction curve

```{r, fig.height = 4.2, fig.width=6, fig.align='center'}
b <- coef(d.logit)
Phat <- function(x) {exp(b[1]+x*b[2])/(1+exp(b[1]+x*b[2]))}
plot(Phat, xlim = c(0,10), xlab="Status of Driver", ylab="Probability of Yielding")
```

## Prediction and data

```{r, fig.height=3, fig.width=5, fig.align='center'}
library(ggplot2)
ggplot(drivers, aes(x = Status, y = as.numeric(Behavior)-1)) +
  geom_point(position=position_jitter(height=0.04, width=0.08)) + 
  ylab("Probability of Yielding") + stat_function(fun=Phat, color="blue", size=1.5)
```

## Prediction and data

```{r, echo=FALSE}
library(ggplot2)
ggplot(drivers, aes(x = Status, y = as.numeric(Behavior)-1)) +
  geom_point(position=position_jitter(height=0.04, width=0.08)) + 
  ylab("Probability of Yielding") + stat_function(fun=Phat, color="blue", size=1.5)
```

## Plot using base r graphics

```{r, fig.height=3, fig.width=5, fig.align='center'}
pred.df <- data.frame(Status = seq(min(drivers$Status), max(drivers$Status), len=100))
pred.df$phat <- predict(d.logit, newdata = pred.df, type="response")
plot(as.numeric(Behavior)-1 ~ Status, data = drivers)
lines(phat ~ Status, data = pred.df, col="blue", lwd = 1.5)
```


# Review of Simulation-Based Tests: Explanatory and Response scenarios

## Two proportion applet

The [Two-Proportion Applet](http://www.rossmanchance.com/applets/ChiSqShuffle.html?hideExtras=2) is preloaded with the Yawning data. 

- Explanatory: Whether or not the subject saw a "yawn seed." (Categorical)
- Response: Whether or not the subject yawned. 

## Two mean applet

The [Two Mean Applet](http://www.rossmanchance.com/applets/AnovaShuffle.htm?hideExtras=2) has the preloaded Sleep Deprivation data.

- Explanatory: Sleep Deprived? (Y/N) (categorical)
- Response: Improvement score on a test. (quantitative)

## Paired data applet

The [Matched Pairs Applet](http://www.rossmanchance.com/applets/MatchedPairs/MatchedPairs.htm?hideExtras=1) has the preloaded Rounding First Base data.

- Explanatory: Two different running strategies (categorical)
- Response: Time to second base (quantitative)

## Multiple proportions applet

The [Multiple Proportions Applet](http://www.rossmanchance.com/applets/ChiSqShuffle.html?hideExtras=1) has the preloaded Organ Donor data.

- Explanatory: Type of recruitment strategy (neutral, opt-in, or opt-out). (categorical) 
- Response: Whether or not the subject chose to donate. (categorical)

## Multiple Means applet

The [Multiple Means Applet](http://www.rossmanchance.com/applets/AnovaShuffle.htm) has the preloaded data. Check *Show Shuffle Options*, and do one shuffle. Observe what happens.

- Explanatory: Type of treatment (A,B,C,D,E) (categorical)
- Response: Strength (quantitative)

## Correlation/Regression

The [Correlation/Regression Applet](http://www.rossmanchance.com/applets/RegShuffle.htm?hideExtras=2) has the preloaded Footprint data.

- Explanatory: foot length (quantitative)
- Response: height (quantitative)

# verbal vs visual memory

## Count the passes

<iframe width="560" height="315" src="https://www.youtube-nocookie.com/embed/vJG698U2Mvo?start=4" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Experiment

- Subjects were given a statement and told to write a justification of its claim. Some subjects had to justify *concrete* claims, while others were given *abstract* claims. Their writing was then assigned a quality rating.
- At the same time, subjects performed an imagery memory task, one involving pictures (*visual*), and the other involving syllables (*verbal*).
- Does the type of claim matter? (concrete/abstract)
- Does the type of imagery matter? (verbal/visual)

## Comparing two groups

- We could compare the mean ratings of the two types of claims.
- Or we could compare the mean ratings of the two imagery groups.

What test would we be using?

## Data 

```{r}
ratings <- read.table("http://math.westmont.edu/ma5/ratings.txt", header = TRUE)
ratings
```

## Summary of `ratings` data

```{r}
summary(ratings)
```


## Two-sample t-test: claims

```{r}
t.test(Rating ~ Claim, data=ratings)
```


## Two-sample t-test: Imagery

```{r}
t.test(Rating ~ Imagery, data=ratings)
```

## Multiple comparisons problem

>- When you have a large number of variables, beware of doing several pairwise tests.
>- Each test will have a chance of making a Type I error (false alarm).
>- Multiple comparisons compound the chances of making a Type I error.
>- Given enough tests to try, one will probably be significant (p-value < 0.05) just by chance, even when the null hypothesis of no association is true.

## Relating the effects of two variables

- The researchers were not only interested in which task was harder.
- They were interested if the type of imagery in the memory task interfered with paragraph writing differently, for concrete and abstract claims.
- In this situation, they were looking for an *interaction* between *imagery* and *claim*, as measured by *rating*.

## Simulation using applet

The data for this study is preloaded into the [Two-variable ANOVA applet](http://www.rossmanchance.com/applets/twoway.html).

1. Click on *Show Means* and **record** the observed *Difference in Differences*, which will be our statistic.

2. Complete the sentence to interpret the Difference in Differences in context: Both groups, Concrete and Abstract, got lower ratings using the _______ imagery, but the ______ group showed a greater effect, by ______ points. 

## Interaction plot

3. The *interaction plot* shows the means and illustrates the Difference in Differences. If the effect of the memory task was the same on both groups, what would the lines look like? **Record** a sketch showing what it would look like if the Difference in Differences were equal to zero.

## Do one simulation

4. Click on *Show Shuffle Options*, and click the *Data* radio button. Do one or two shuffles. Look at the data and **record**  what variable is getting shuffled, and whether it is the explanatory or response. You may have to enlarge the data panels to be able to see all the data.

5. Now click on the *Graphs* radio button, and do a few more shuffles. How would you characterizes the randomly generated interaction plots compared to the observed interaction plot?

## Compute a p-value

We will test the null hypothesis that the population Difference in Differences is zero, versus the alternative that it is not equal to zero.

6. Do at least 1000 shuffles, and **record** a p-value and how strong the evidence is against the null hypothesis. What do you think we have shown, in the context of this study?

## Interaction effects

- The ratings of Concrete definitions were higher than Abstract definitions.
- Both groups did worse during the Verbal task.
- However, the effect was stronger on the Concrete group. 
- So there is a significant *interaction* between the two variables. The value of one variable influences the effect of the other.
- If there is no significant interaction, you could test the ratings for all the Concrete vs. all the Abstract (or all Visual vs. all Verbal)
- If there is an interaction, you should compare means for all four combinations of categories.

## Find the Theory-based test

We have two explanatory variables, Imagery (Visual/Verbal) and Claim (Concrete/Abstract) and one response variable, Rating.

7. Use the [UCLA guide](https://stats.idre.ucla.edu/other/mult-pkg/whatstat/) to determine which test to use. **Record** the name of this test.

## Interaction Plots in R

```{r fig.height=4, fig.width=7, fig.align='center'}
ratings$Imagery <- factor(ratings$Imagery,levels(ratings$Imagery)[c(2,1)]) # reorder factors
interaction.plot(ratings$Imagery, ratings$Claim, ratings$Rating)
```

## Boxplots for multiple categories

```{r fig.height=4.2, fig.width=6, fig.align='center'}
ggplot(ratings, aes(x=Imagery, y=Rating, fill=Claim)) + geom_boxplot()
```


## Two-way ANOVA in R

We use an asterisk * instead of a plus + in the formula, because we want to test for interaction effects.

```{r}
r.aov <- aov(Rating ~ Claim * Imagery, data = ratings)
anova(r.aov)
```

## Followup-test

```{r}
TukeyHSD(r.aov)
```








