---
title: "Section 10.3"
date: "November 22, 2019"
output: 
  word_document:
  revealjs::revealjs_presentation:
    fig_width: 14
    fig_height: 7
    self_contained: true
    theme: black
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 600
---

# Section 10.3: Least Squares Regression

## Regression Line

- If we decide an association is linear, it’s helpful to develop a mathematical model of that association. 
- Helps make predictions about the response variable. 
- The least-squares regression line is the most common way of doing this.  

## Review: Linear Functions

>- A *linear function* has an equation of the form $f(x) = a + bx$.
>- The graph a function is the set of all points of the form $(x, f(x))$.
>- The graph of a linear function is a straight line.
>- The slope of this linear function is the constant $b$. The slope is the vertical change over the horizontal change (rise/run).
>- The $y$-intercept of this function is the constant $a$.

# Regression Line

## {data-background="https://upload.wikimedia.org/wikipedia/commons/thumb/4/40/Cptvdisplay.jpg/1200px-Cptvdisplay.jpg" data-background-size="contain"}

## Line of Best Fit 

- Unless the points form a perfect line, there won’t be a single line that goes through every point.  
- We want a line that gets as close as possible to all the points.
- *Note:* Correlation $\neq$ Causation

```{r, echo=FALSE, fig.width=4, fig.height=3, fig.align='center'}
library(ggplot2)
tvlife <- read.table("http://www.math.hope.edu/isi/data/chap10/TVLife.txt", header = TRUE)
ggplot(tvlife, aes(x=TVsperK, y=LifeExpectancy)) + geom_point() +geom_smooth(method=lm, se=FALSE)
```

## Residuals

- The vertical distances between the data points and the line are called **residuals**. 
- The **regression line** minimizes the sum of the squares of the residuals.

```{r, echo=FALSE, fig.width=4, fig.height=3, fig.align='center', warning=FALSE}
library(broom)
mod <- lm(tvlife$LifeExpectancy ~ tvlife$TVsperK)
df <- augment(mod)
ggplot(df, aes(x=tvlife.TVsperK, y=tvlife.LifeExpectancy)) + geom_point() +geom_smooth(method=lm, se=FALSE) + geom_segment(aes(xend = tvlife.TVsperK, yend=.fitted))
```

## {data-background="http://imgs.xkcd.com/comics/correlation.png" data-background-size="contain"}


# Exploration 10.1, Revisited

## {data-background="https://i.pinimg.com/originals/55/ba/1e/55ba1e76220081d02f1a2405fe3d6186.jpg" data-background-size="contain"}

## The Delboeuf Illusion

- Researchers gathered data to investigate the claim that dinner plates are growing
- American dinner plates sold on eBay on March 30, 2010 (Van Ittersum and Wansink, 2011).
- Year manufactured and diameter are given.

```{r, echo=FALSE}
plates <- read.table("http://www.math.hope.edu/isi/data/chap10/PlateSize.txt", header = TRUE)
plates
```

## Plate Size vs. Year

- Both year (explanatory variable) and diameter in inches (response variable) are quantitative.
- Each dot represents one plate in this scatterplot.

```{r, echo=FALSE, fig.width=4, fig.height=3, fig.align='center'}
ggplot(plates, aes(x=year, y=size)) + geom_point()
```

## Plate Size vs. Year

- Regression line: $\hat{y} = -14.8 + 0.0128x$

```{r, echo=FALSE, fig.width=4, fig.height=3, fig.align='center'}
ggplot(plates, aes(x=year, y=size)) + geom_point() + geom_smooth(method=lm, se=FALSE)
```

## Formulas

\[ 
\begin{align}
\hat{y} &= a + bx && \text{regression line} \\
b &= r \frac{s_y}{s_x} && \text{sample slope} \\
a &= \bar{y} - b\bar{x} && \text{sample $y$-intercept} \\
\end{align}
\]

- For a given value of $x$, $\hat{y}$ is the corresponding *predicted* value.
- The distance $y - \hat{y}$ (actual - predicted) is a *residual*.

## Residuals

- The distance $y - \hat{y}$ (actual - predicted) is a *residual*.
- The residuals can be represented by vertical line segments.
- The regression line minimizes the residuals.

```{r, echo=FALSE, fig.width=4, fig.height=3, fig.align='center', warning=FALSE}
library(broom)
mod2 <- lm(plates$size ~ plates$year)
df2 <- augment(mod2)
ggplot(df2, aes(x=plates.year, y=plates.size)) + geom_point() +geom_smooth(method=lm, se=FALSE) + geom_segment(aes(xend = plates.year, yend=.fitted))
```


## Predictions

> - The regression line for plate size as a function of year is $\hat{y} = -14.8 + 0.0128x$.
> - We don't have data for the year 2000, but we can **predict** that a plate from the year 2000 has size $\hat{y} = -14.8 + 0.0128(2000) = 10.8$. 
> - Similarly, a 2001 plate has a predicted size of $-14.8 + 0.0128(2001) = 10.8128$.
> - Notice that our model predicts that plates grow by 0.0128 inches per year. This is what the slope tells you.

## Slope

- Slope is the predicted change in the response variable for one-unit change in the explanatory variable.
- Both the slope and the correlation coefficient for this study were positive.
    + The slope is 0.0128
    + The correlation is 0.604 
- The slope and correlation coefficient will always have the same sign. ($b = r \frac{s_y}{s_x}$)

## Interpolation vs Extrapolation

- The regression equation $\hat{y} = -14.8 + 0.0128x$ should only be used to predict sizes for about 1950 to 2010 (the range of our data). This is called **interpolation**.
- Predicting values for the response variable for values of the explanatory variable that are outside of the range of the original data is called **extrapolation**.
    + Our model predicts that in the year 2500, Americans will be eating off of plates that are 17.2 inches in diameter.
- **Typically, you should avoid extrapolation.**


## {data-background="http://imgs.xkcd.com/comics/extrapolating.png" data-background-size="contain"}

## {data-background="http://math.westmont.edu/img/rvsxkcd.jpg" data-background-size="contain"}

## Coefficient of Determination

- While the intercept and slope have meaning in the context of year and diameter, remember that  the correlation does not. It is just 0.604, with no units.
- However, the square of the correlation (coefficient of determination or $r^2$) does have meaning.
- $r^2  = (0.604)^2 \approx 0.365 = 36.5\%$
- **Meaning:** 36.5% of the variation in plate size (or the response variable) can be explained by its linear association with the year (the explanatory variable).

## Overview of applet

Paste the [PlateSize data set](http://www.math.hope.edu/isi/data/chap10/PlateSize.txt) into the [Correlation/Regression applet](http://www.rossmanchance.com/applets/RegShuffle.htm?hideExtras=2).

- Create scatterplot
- Find correlation
- Find regression equation
- Find coefficient of determination

# Exploration 10.3

## Crime Scene Footprints {data-background="https://i.pinimg.com/originals/be/15/74/be1574970912db22164d07a5bfa9a76d.jpg" data-background-size="contain"}

## Preview

Can a footprint taken at the scene of a crime help to predict the height of the criminal? In other words, is there an association between height and foot length? To investigate this question, a sample of 20 statistics students measured their height (in inches) and their foot length (in centimeters).

- Observational units: The 20 stats students
- Explanatory variable: Foot length (quantitative)
- Response variable: Height (quantitative)
- Data indicates a moderate to strong positive linear correlation.

## **Don't click on "show regression line" yet!!**

Please **resist the temptation** to click on "show regression line" until you have completed Question 3 and entered your guess in the form.

## Guessing game

*Before you calculate the regression line,* open the preloaded foot data in the [Correlation/Regression applet](http://www.rossmanchance.com/applets/RegShuffle.htm?hideExtras=2).

1. Check the *Show Movable Line* box to add a blue line to the scatterplot. Move the line until you believe your line "best" summarizes the relationship between height and foot length for these data. **Record** the resulting equation for your line (using traditional statistical notation).
    
## Minimize residuals

*A residual is the difference between an observed response and the corresponding prediction made by the least squares regression line (residual = observed - predicted).*

2. Check the *Show Residuals* box to visually represent these residuals for your blue "guess" line. **Record** the *SAE* value (Sum of Absolute Errors) and the *SSE* (Sum of Squared Errors).

3. Now continue to adjust your blue line until you think you have minimized the sum of the squared residuals. **Record** your new equation and your new SSE value. **Operator:** [Enter your SSE in this form](https://goo.gl/forms/m3Womm0twhsoMCuY2).

## Regression line

4. *Now* check the *Show Regression Line* box and **record** the equation of the least squares regression line.  Also **record** the SSE. How does the SSE compare with the value you found in #3?

5. Use the least squares regression line to predict the height of someone whose foot length is 29 cm. **Record** your prediction.

6. Interpret the slope in context: The slope of the regression line predicting height based on foot length is __________, meaning that for every additional _______ cm increase in foot length, the predicted height increases by _________ inches.

## Compute a residual

> A **residual** is the difference between an observed response and the corresponding prediction made by the least squares regression line (residual = observed - predicted = $y - \hat{y}$).

7. **Record** the residual for the person whose foot length is 32 centimeters and height is 74 inches.

## Influential observations 

8. Uncheck *Show Movable Line* and check the *Move Observations* box.
    + Choose a point in the middle of the scatterplot and move it up and down.
    + Now choose a point on the far right of the scatterplot and move it up and down.
    
Which of these two observations is more influential? Generally speaking, where do influential observations tend to be found? **Record** your answer.

## Extrapolation

9. A size 0 infant shoe is about 8cm long. **Record** the regression line's prediction for the height of an infant with this shoe size?  Why does this model produce such an unrealistic prediction?

## Coefficient of Determination

10. Check the *Correlation coefficient* box and the *R-squared* box. **Record** $r$ and $r^2$. Complete the following sentence:

The coefficient of determination is __________%; this means that __________% of the variation in people's ___________________ is attributable to the linear association with ______________________________. 


## {data-background="http://math.westmont.edu/img/corvscause.jpg" data-background-size="contain"}
