---
title: "Section 2.1"
date: "September 9, 2019"
output: 
  word_document:
  revealjs::revealjs_presentation:
    fig_width: 14
    fig_height: 7
    self_contained: true
    theme: black
    highlight: zenburn
    css: slidesdjh.css
    center: true
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 600
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(comment = NA)
```

# Section 2.1: Sampling from a Finite Population

## Last time: Heads or Tails

>- We used our own class survey to test whether the assertion of Steve Campbell was plausible, namely, that people pick heads 70% of the time when they're asked to predict the outcome of a coin toss.
>- In our sample, 33/64 picked heads.
>- We found that $H_0$ (Campbell's assertion) was not plausible (small p-value), based on our sample.
>- Two possibilities:
>    - Campbell is wrong.
>    - Campbell is right, but our *sample* is somehow not representative of all people.


# Generalization: How broadly do our results apply?

## How broadly do our results apply?

- We are usually interested in a *population* (e.g., all people, all college students)
- We usually collect data from a *sample* (e.g., people in our dorm)

What does our statistical analysis tell us about the population of interest?

## Dewey Defeats Truman

One of the most famous instances of nonrepresentative sampling involved the polls prior to the 1948 US presidential election.

<iframe width="560" height="315" src="https://www.youtube-nocookie.com/embed/5BK8HxzIWs0?rel=0" frameborder="0" allow="autoplay; encrypted-media" allowfullscreen></iframe>

## What happened?

The newspaper thought Dewey would win because of poll results. [However](http://www.oxfordmathcenter.com/drupal7/node/251),

- They were relying on opinion polls done weeks before the election.
- Polls were conducted over the telephone. (In 1948, who had telephones?)
- Pollsters were allowed to choose whom to ask, as long as they met demographic quotas. (convenience sample)

## Parameters vs. Statistics

The **population** is the entire collection of observational units we are interested in, and the **sample** is a subset of the population on which we record data. Numerical summaries about a population are called **parameters**, and numerical summaries calculated from a sample are called **statistics**. For example:

- Population: all US voters in 1948.
- Parameter: the proportion of all voters who support Truman
- Sample: the people that the pollsters selected.
- Statistic: the proportion of the sample who support Truman

# Exploration 2.1A: Sampling Words

## Preview

1. For the parts that follow, your group should use the **Operator's** sample of 10 words from the Gettysburg Address that was made in the Preview assignment. If the Operator doesn't have that sample, use someone else's sample.

- *Population:* all of the words in the passage.
- *Sample:* the 10 words you selected.

(Note: this is a "toy example.")

## Make a table

<div class="column-left">
2. Write down each word from your sample and then indicate the length of the word (number of letters) and whether or not the word contains at least one letter e.

</div>

<div class="column-right">

| Word | Length | Contains e? |
| ---- | ------ | ----------- |
| 1 | | |
| 2 | | |
| 3 | | |
| 4 | | |
| 5 | | |
| 6 | | |
| 7 | | |
| 8 | | |
| 9 | | |
| 10 | | |

</div>

## Data tables

A **data table** (or statistical spreadsheet) is a convenient way to store and represent all data values. Typically, in a data table, the *rows correspond to observational units,* columns represent variables, and the data are the table entries.

## Average word length

5. Calculate the *average length* of the 10 words in your sample. Is this number a parameter or a statistic? Explain how you know. What symbol would you use to refer to this value? **Recorder:** Record the average word length, using appropriate symbols. **Presenter:** Write this on the board.

6. The average length of the 268 words in the entire speech equals 4.29 letters. Is this number a parameter or a statistic? Explain how you know. What symbol would you use to refer to this value?

## Proportion of e-words

7. Calculate the *proportion of words in your sample that contain at least one e.* Is this number a parameter or a statistic? Explain how you know. What symbol would you use to refer to this value? **Recorder:** Record the proportion of e-words, using appropriate symbols.

8. The proportion of all words in the entire speech that contain at least one letter e is $125/268 \approx 0.47$. Is this number a parameter or a statistic? Explain how you know. What symbol would you use to refer to this value?

## Dotplot of word lengths

10. Let's make a dot plot of the average word lengths for each group's sample. 
    a. How do we label the axis of this dotplot appropriately?
    b. Explain what each dot on the dotplot represents by filling in the following sentence. Each dot on the dotplot is a single measurement of the value of ___ on a single ___.
    c. Do you think your samples are representative of the population? How are you deciding?
    
## Data for our class

```{r}
samples <- readLines("http://math.westmont.edu/img/samplewords19.txt")
samples
```

## Dotplot for our class

```{r, warning=FALSE, message=FALSE}
samples <- readLines("http://math.westmont.edu/img/samplewords19.txt")
n <- length(samples)
aveWordLength <- numeric(n)
eWordProp <- numeric(n)
sdWordLength <- numeric(n)
for(i in 1:n) {
  words <- strsplit(samples[i]," ")[[1]]
  wordlengths <- nchar(words)
  aveWordLength[i] <- mean(wordlengths)
  sdWordLength[i] <- sd(wordlengths)
  numEs <- sum(grepl("e", words))
  eWordProp[i] <- numEs/length(words)
}
classData <- data.frame(aveWordLength, eWordProp, sdWordLength)
#print(samples)
#print(classData)
library(ggplot2)
ggplot(classData, aes(x = aveWordLength)) + geom_dotplot(dotsize=0.8) + scale_y_continuous(NULL, breaks = NULL)
```

## Bias?

11. Let's compare our sample statistics to the population parameters. ($\mu = 4.29$, $\pi = 0.47$)
```{r}
cat("Number of samples: ", nrow(classData), "\n")
cat("Average word length: ", mean(classData$aveWordLength), "\n")
cat("Number of samples with ave word length > 4.29: ", sum(classData$aveWordLength > 4.29), "\n")
cat("Proportion of samples with ave word length > 4.29: ", sum(classData$aveWordLength > 4.29)/n, "\n\n")
cat("Number of samples with proportion e words > 0.47: ", sum(classData$eWordProp>0.47), "\n")
cat("Proportion of samples with proportion e words > 0.47: ", sum(classData$eWordProp>0.47)/n, "\n")
```

## Biased Sampling Method

- Bias is a property of a sampling method, not a property of an individual sample. 

- A sampling method is **biased** if, when using that sampling method, statistics from different samples consistently overestimate or consistently underestimate the population parameter of interest.

## Reason for bias?

13. Discuss why we might have expected this sampling method (asking you to quickly pick 10 representative words) to be biased. 

14. Do you think asking each of you to take 20 words instead of 10 words would have helped with this issue? Explain.

## Pointing while blindfolded?

What if you were to close your eyes and point blindly at the page with a pencil 10 times, taking for your sample the 10 words on which your pencil lands.

15. Do you think this sampling method is likely to be biased? If so, in which direction? Explain.

A sample is only useful to us if the data we collect on our sample are similar to the results we would find in the entire population. In this sense, we say the sample is *representative* of the population. 

## Simple Random Sample

> A **simple random sample** (SRS) ensures that every sample of size $n$ is equally likely to be the sample selected from the population. In particular, each observational unit has the same chance of being selected as every other observational unit.

## Collect a simple random sample

17. **Operator:** Go to the [Random Numbers Applet](http://www.rossmanchance.com/applets/RandomGen/GenRandom01.htm). Specify that you want *one replication* with *5 Numbers per replication* in the *range from 1 to 268*, *without replacement*. Press Generate to view the 5 random numbers. **Record** your five random numbers in the first column of a table, like the one below. Using your randomly generated values, look up the corresponding word from the sampling frame in the book (pages 112-113). 

| Random Number | Word | Length | Contains e? |
| ---- | ------ | ----------- | ----------- |
| &nbsp; | | | |
|  | | | |
|  | | | |
|  | | | |
|  | | | |


## Sampling Frame

```
ID	Word	
1	Four
2	score	
3	and
4	seven	
5	years	
6	ago
7	our
8	fathers	
9	brought
10	forth
11	upon
12	this
13	continent	
14	a
15	new	
16	nation	
17	conceived	
18	in
19	liberty	
20	and
21	dedicated	
22	to
23	the	
24	proposition	
25	that
26	all
27	men	
28	are	
29	created	
30	equal	
31	Now
32	we	
33	are	
34	engaged	
35	in
36	a
37	great	
38	civil
39	war	
40	testing	
41	whether	
42	that
43	nation	
44	or
45	any
46	nation	
47	so
48	conceived	
49	and
50	so
51	dedicated	
52	can
53	long
54	endure	
55	We	
56	are	
57	met	
58	on
59	a
60	great	
61	battlefield	
62	of
63	that
64	war	
65	We	
66	have	
67	come	
68	to
69	dedicate	
70	a
71	portion	
72	of
73	that
74	field	
75	as
76	a
77	final
78	resting	
79	place	
80	for
81	those	
82	who
83	here	
84	gave	
85	their	
86	lives	
87	that
88	that
89	nation	
90	might
91	live	
92	It
93	is
94	altogether	
95	fitting
96	and
97	proper	
98	that
99	we	
100	should
101	do
102	this
103	But
104	in
105	a
106	larger	
107	sense	
108	we	
109	cannot
110	dedicate	
111	we	
112	cannot
113	consecrate	
114	we	
115	cannot
116	hallow
117	this
118	ground	
119	The	
120	brave	
121	men	
122	living
123	and
124	dead	
125	who
126	struggled	
127	here	
128	have	
129	consecrated	
130	it
131	far
132	above	
133	our
134	poor
135	power	
136	to
137	add
138	or
139	detract	
140	The	
141	world	
142	will
143	little	
144	note	
145	nor
146	long
147	remember	
148	what
149	we	
150	say
151	here	
152	but
153	it
154	can
155	never	
156	forget	
157	what
158	they	
159	did
160	here	
161	It
162	is
163	for
164	us
165	the	
166	living	
167	rather	
168	to
169	be	
170	dedicated	
171	here	
172	to
173	the	
174	unfinished	
175	work	
176	which
177	they	
178	who
179	fought
180	here	
181	have	
182	thus
183	far
184	so
185	nobly
186	advanced	
187	It
188	is
189	rather	
190	for
191	us
192	to
193	be	
194	here	
195	dedicated	
196	to
197	the	
198	great	
199	task	
200	remaining	
201	before	
202	us
203	that
204	from
205	these	
206	honored	
207	dead	
208	we	
209	take	
210	increased	
211	devotion	
212	to
213	that
214	cause	
215	for
216	which
217	they	
218	gave	
219	the	
220	last
221	full
222	measure	
223	of
224	devotion	
225	that
226	we	
227	here	
228	highly
229	resolve	
230	that
231	these	
232	dead	
233	shall
234	not
235	have	
236	died	
237	in
238	vain	
239	that
240	this
241	nation	
242	under	
243	God	
244	shall
245	have	
246	a
247	new	
248	birth	
249	of
250	freedom	
251	and
252	that
253	government	
254	of
255	the	
256	people	
257	by
258	the	
259	people	
260	for
261	the	
262	people	
263	shall
264	not
265	perish	
266	from
267	the	
268	earth	
```

## Calculate statistics

18. Let's examine your sample and those of your classmates. Calculate the *average word length* and *proportion of e-words* in your random sample. **Presenter:** Write these numbers on the board.

## Bias?

19. Let's compare your sample statistics to the population parameters.
    a. How many groups obtained a sample average word length larger than the population average (4.29 letters)?
    b. How many groups obtained a sample proportion of e-words larger than the population proportion (0.47)? 
    c. What do these answers reveal about whether simple random sampling is an unbiased sampling method? 

## Simulation

20. Open the [Sampling Words applet](http://www.rossmanchance.com/applets/OneSample.html?population=gettysburg). For now, use the pull-down menu to select the *length* variable. 

21. Check the *Show Sampling Options* box. Specify 5 in the *Sample Size* box and press the *Draw Samples* button. You will see five blue dots within the population distribution representing the five lengths that you have sampled, which are also displayed in a dotplot in the middle panel. The blue triangle indicates the value of this sample's mean, which is then added to the graph on the right. 

## Do 1000 samples total

22. Press the Draw Samples button four more times (for a total of five samples). Notice how the samples vary and how the sample mean varies from sample to sample.Change the Number of Samples to 995 (for a total of 1,000 samples). Press the Draw Samples button. (This should generate a dotplot of 1,000 values of the sample mean for 1,000 different simple random samples, each consisting of 5 words, selected randomly from the population of all 268 words in the Gettysburg Address.) 

**Recorder:** Record the mean and standard deviation of the distribution of sample means.

## Effect of sample size?

24. Now change the Sample Size from 5 to 20 and the Number of Samples to 1000. Press the Draw Samples button. How do these two distributions compare? **Recorder:** Write down a feature that is similar and a feature that is different.

25. Did changing the sample size change the center of the distribution? If we used a biased sampling method, would increasing the sample size remove the bias? 

## Summary

**Presenter:** Write a sentence on the board describing a concept we explored today. (Use some of these terms: "SRS", "sample", "bias", "population", "statistic", "parameter", "mean", "standard deviation")
