---
title: "Section 3.3"
date: "September 27, 2019"
output: 
  word_document:
  revealjs::revealjs_presentation:
    fig_width: 14
    fig_height: 7
    self_contained: true
    theme: black
    highlight: zenburn
    css: slidesdjh.css
    center: true
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 600
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(comment = NA)
```

# Section 3.3: 2SD and Theory-based Confidence Intervals for a Single Mean

## Introduction

- Last time, we saw shortcuts for constructing confidence intervals for a population proportion $\pi$.
    - The variables were *categorical*, so we summarized the data with $\hat{p}$.
- Today, we are going to use similar shortcuts to construct confidence intervals for a population mean $\mu$.
    - The variables will be *quantitative*, we will summarize the data with $\bar{x}$ and $s$.
    
## Example: Used Cars {data-background="http://classiccardb.com/uploads/postfotos/1982-honda-civic-sedan-classic-all-original-survivor-w-59k-miles-2.jpg"}

**Question:** How much does a used Honda Civic cost?

## Used Car Data

The following histogram displays data for the selling price of 102 Honda Civics that were listed for sale on the Internet.

```{r, fig.height = 5}
library(ggplot2)
carPrices <- read.table("http://www.isi-stats.com/isi/data/chap3/UsedCars.txt", header=TRUE)
ggplot(data = carPrices, aes(Price)) + geom_histogram(binwidth = 2000)
```

## Used Cars

- While we should be cautious about our sample being representative of the population, let’s treat it as such.
- The sample mean is $\bar{x} = \$13,292$.
- We don’t presume that $\mu = \$13,292$, but it should be "close."
- To determine how close, we can construct a confidence interval.

## Confidence Intervals

- Remember the basic form of a confidence interval is:  
$$
\text{statistic} \pm (\text{multiplier}) \times (\text{SD of statistic})
$$

- In our case, the statistic is $\bar{x}$, so we write our 2SD confidence interval as:
$$
\bar{x} \pm 2 \times (\text{SD of } \bar{x})
$$

- We need a way to estimate SD of $\bar{x}$.


## Confidence intervals

- *Important:* The SD of $\bar{x}$ and the SD of our sample ($s = \$4,535$) are not the same.
- There is more variability in the data (the car-to-car variability) than in sample means.
- We can approximate the variability in the sample means as $s / \sqrt{n}$.
- So we can write a 2SD confidence interval as:
$$
\bar{x} \pm 2 \times \frac{s}{\sqrt{n}}
$$

- This method gives a rough approximation for a 95% CI.

## Theory-based CI for $\mu$

- A theory-based confidence interval is quite similar except it uses a multiplier that is based on a t-distribution and is dependent on the sample size and confidence level. (We will let the applet figure all this out for us.) 
- For theory-based confidence interval for a population mean (called a one-sample t-interval) to be *valid*, the sample distribution should be symmetric or we need to have a sample size of at least 20 and the sample distribution should not be strongly skewed.

## Theory-based CI for $\mu$

The confidence interval for a population mean has the form: 
$$
\bar{x} \pm (\text{multiplier}) \times \frac{s}{\sqrt{n}}
$$

- $\bar{x}$ is the center (our best estimate for the population mean).
- $s/\sqrt{n}$ is the predicted value of the standard deviation of the null distribution (also called the standard error).
- $\text{multiplier}$ is a number that is dependent upon the confidence level ($\approx$ 2 for 95% confidence intervals)

## Data

```{r, echo=TRUE}
carPrices
```

## Summary Statistics

```{r, echo=TRUE}
mean(carPrices$Price)
sd(carPrices$Price)
```

2SD interval: $13292 \pm 2 \times \frac{4534.568}{\sqrt{102}} \approx (12394, 14190)$

## Theory-based inference applet

We can just paste the [Used Car data](http://www.isi-stats.com/isi/data/chap3/UsedCars.txt) into the [Theory-based inference applet](http://www.rossmanchance.com/applets/TBIA.html?hideExtras=1).

## Theory-based interval in R

```{r, echo=TRUE}
t.test(carPrices$Price)
```

- Compare to 2SD interval: (12394, 14190)
- 2SD interval is a little wider, but close (3 sig figs)

# Exploration 3.3: Sleepless Nights? (continued) {data-background="https://thumbs.gfycat.com/ShowyEvenFreshwatereel-small.gif"}

## Sleep data
 
1. In Exploration 2.2, we collected data on how many hours of sleep we got. Use the [Descriptive Statistics applet](http://www.rossmanchance.com/applets/Dotplot.html?hideExtras=1) to get $n$, $\bar{x}$, and $s$ for [our class data](https://math.westmont.edu/ma5/classSleep.txt). **Record** these values.

2. Based on this data, what is the best estimate for $\mu$, the average amount of sleep Westmont students get?

3. It was tricky to generate a null distribution to estimate $\text{SD}_\text{null}$, the standard deviation of sample means. Instead, **record** the standard error, $s/\sqrt{n}$.

## Validity Conditions

> The theory-based interval for a population mean (called a *one-sample t-interval*) requires that the quantitative variable should have a symmetric distribution or you should have at least 20 observations and the sample distribution should not be strongly skewed.

4. In Exploration 2.2, we decided that these validity conditions hold for our data.

## 2SD Confidence interval

5. Compute a 95% 2SD confidence interval for $\mu$, using the standard error you recorded in #3. **Record** this interval.

6. Complete this sentence, **recording** what goes in the blanks. We are 95% confident that _____ is between _____ and _____.

## Theory-based inference applet

7. We can also use the [Theory-Based Inference applet](http://www.rossmanchance.com/applets/TBIA.html?hideExtras=1) to get confidence intervals. Use the *One Mean* scenario, check the boxes for *Paste Data* and *Includes header*, *Clear* the sample data, and paste in [our class data](https://math.westmont.edu/ma5/classSleep.txt). **Record** the 95% theory-based t-interval from the applet. Compare it to the 2SD interval that you calculated in #5-#6.

8. **Record** a 90% confidence interval also.

9. Which interval is wider, the 95% or the 90%? **Presenter:** Be ready to explain why it makes sense that this interval is wider.

## CI's and Hypothesis tests

Use your CI's from #7 and #8 to answer the following.

11. At a signficance level of $\alpha = 0.05$, would you reject $H_0 : \mu = 6.5$ in favor of the alternative $H_a : \mu \neq 6.5$?

12. At a signficance level of $\alpha = 0.10$, would you reject $H_0 : \mu = 6.5$ in favor of the alternative $H_a : \mu \neq 6.5$?

**Record** your answers. **Presenter:** Write a sentence or diagram on the board explaining how you can use a confidence interval to determine the result of a hypothesis test.

## One-mean confidence intervals in R

```{r, echo=TRUE}
sleepDF <- read.table("https://math.westmont.edu/ma5/classSleep.txt", header = TRUE)
t.test(sleepDF$SleepHours)
```

This is more accurate than the applet.

## One-mean confidence intervals in R

You can change the confidence level:

```{r, echo=TRUE}
t.test(sleepDF$SleepHours, conf.level = 0.90)
```

## One-mean p-values in R

You can set the null value for a two-sided hypothesis test:

```{r, echo=TRUE}
t.test(sleepDF$SleepHours, conf.level = 0.90, mu = 6.5)
```
