---
title: "Section 4.1"
date: "October 2, 2019"
output: 
  word_document:
  revealjs::revealjs_presentation:
    fig_width: 14
    fig_height: 7
    self_contained: true
    theme: black
    highlight: zenburn
    transition: slide
    css: slidesdjh.css
    center: true
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 600
---

# Chapter 4

## Big idea of chapter 4

* Previously research questions focused on a single statistic.
    + What proportion of the time did Harley choose the right bowl?  
    + What is the average hours slept by Westmont Students? 

* We will now start to focus on research questions comparing two groups.  
    + Are smokers more likely than nonsmokers to have lung cancer? 
    + Are children who used night lights as infants more likely to need glasses than those who didn’t use night lights?
    
# 4.1: Association and Confounding

## Types of variables

<span style="color:lightgreen">
When two variables are involved in a study, they are often classified as **explanatory** and **response**.
</span>

> * Explanatory variable (Independent, Predictor)
>     + The variable we think is “explaining” the change in the response variable. (Many times, this is the variable the researchers are manipulating.)
> * Response variable (Dependent)
>     + The variable we think is being impacted or changed by the explanatory variable.  

## Examples

> * Choose the explanatory and response variable:
>     + Smoking and lung cancer 
>     + Heart disease and diet
>     + Height and weight
> * <span style="color:yellow">Sometimes there is a clear distinction between explanatory and response variables and sometimes there isn’t.</span>

# Night Lights and Nearsightedness

## Near-sightedness often develops in childhood {data-background="https://images-na.ssl-images-amazon.com/images/I/31vlNtv87GL.jpg"}

>* <span style="color:orchid">Recent studies looked to see if there is an association between near-sightedness and night light use with infants</span>
>* <span style="color:orchid">Researchers interviewed parents of 479 children who were outpatients in a pediatric ophthalmology clinic</span>
>* <span style="color:orchid"> Asked whether the child slept with the room light on, with a night light on, or in darkness before age 2</span>
>* <span style="color:orchid"> Children were also separated into two groups: near-sighted or not near-sighted based on the child’s recent eye examination</span>

##  Night Light Data

```{r, echo=FALSE}
library(knitr)
nldata <- data.frame(c(18,154), c(78,154), c(41,34))
colnames(nldata) <- c("Darkness", "Night Light", "Room Light")
nldata$Total <- rowSums(nldata)
nldata[3,] <- colSums(nldata)
row.names(nldata) <- c("Near-sighted", "Not near-sighted", "Total")
kable(nldata, caption = "Night Light Data")
```

<br>

> * This type of table is called a **two-way table** (or "contingency table" or "cross tabulation", or just "crosstab").

##

```{r, echo=FALSE}
kable(nldata, caption = "Night Light Data")
kable(nldata / rep(as.numeric(nldata[3,]), each=3), digits=3, caption="Conditional Column Proportions")
# kable(nldata/nldata$Total, digits= 3, caption="Conditional Row Proportions")
```

## Association

```{r, echo=FALSE}
kable(nldata / rep(as.numeric(nldata[3,]), each=3), digits=3, caption="Conditional Column Proportions")
```

> - Notice that as the light level increases, the percentage of near-sighted children also increases. 
> - We say there is an **association** between near-sightedness and night lights.
> - Two variables are **associated** if the values of one variable provide information (help you predict) the values of the other variable.

## Stacked Bar Plots

```{r, echo=FALSE}
barplot(as.matrix(nldata[1:2,1:3]))
```

Near sighted is dark gray, not-near sighted is light gray.

## Mosaic Plots

```{r, echo=FALSE}
nsmatrix <- matrix(c(18,78,41,154,154,34), byrow = TRUE, nrow=2)
colnames(nsmatrix) <- c("Darkness", "Night Light", "Room Light")
rownames(nsmatrix) <- c("Near-sighted", "Not near-sighted")
nstable <- as.table(nsmatrix)
plot(t(nstable), color=c("purple4","blue"), main="Mosaic Plot of Night Light Data")

```

## Association $\neq$ Causation

- A **confounding variable** is associated to both the explanatory variable and the response variable.
- We say it is **confounding** because its effects on the response cannot be separated from those of the explanatory variable.
- Because of this, we can’t draw cause and effect conclusions when confounding variables are present.


> - Possession of lighters is associated to lung cancer. (Confounding variable: smoker/nonsmoker)
> - <span style="color:yellow">(Confounding variable: eyesight of parents) Nearsighted children have nearsighted parents, and nearsighted parents need lit rooms to function.</span>




# Exploration 4.1: Home Court Disadvantage?

## Home Court advantage

- Crowd distraction (0:15)
- Referee Bias (3:24)   


<iframe width="560" height="315" src="https://www.youtube-nocookie.com/embed/BT55zpjT0v4?rel=0&amp;start=15&amp;end=248" frameborder="0" allow="autoplay; encrypted-media" allowfullscreen></iframe>

## Home Court Disadvantage?

The 2008-9 Oklahoma City Thunder had a win-loss record that was actually worse for home games with a sell-out crowd (3 wins and 15 losses) than for home games without a sell-out crowd (12 wins and 11 losses). 

1. Identify the observational units and variables in this study. 
2. When did the Thunder have a higher winning percentage: in front of a sell-out crowd or a smaller crowd? 
3. Do the two variables appear to be associated?
4. Which would you consider the explanatory variable in this study? Which is the response? (That is, what are the roles of these variables in this study?)

## Data in R

```{r}
okcmatrix <- matrix(c(3,15,12,11), byrow=FALSE, nrow=2)
rownames(okcmatrix) <- c("Wins", "Losses")
colnames(okcmatrix) <- c("Sell-out Crowd", "Smaller Crowd")
okctable <- as.table(okcmatrix)
okctable
prop.table(okctable, 2)
```

## Mosaic Plot of OKC data

```{r, echo=FALSE}
plot(t(okctable), color=c("cyan", "orange"), main="OKC wins and crowd sizes")
```


## Confounding Variables

<span style="color:orchid">
A **confounding variable** is a variable that is related both to the explanatory and to the response variable in such a way that its effects on the response variable cannot be separated from the effects of the explanatory variable.
</span>

5. *Discuss:*  Come up with a plausible confounding variable that would explain why the team would be less likely to win in front of a sell-out crowd than in front of a smaller crowd. **Facilitator:** Get your group to agree on a confounding variable.

## Confounding variable

6. Explain how the variable is confounding--*what is the link between this third variable and the response variable, and what is the link between this third variable and the explanatory variable?* (Hint: Remember that this variable has to be recorded on the observational units: home games for the Thunder.) **Operator:** [Use this form](https://goo.gl/forms/akV1Bu3OlKVXoB1G3) to type up a complete sentence of explanation.

## Crowd size and Opponent records

Of the Thunder's 41 home games, 22 were against teams that won more than half of their games ("strong opponents"). Of these 22 games, 13 were sell-outs. Of the 19 games against "weak opponents"" that won less than half of their games that season, only 5 of those games were sell-outs.

7. **Record** this data in a two-way table. The rows should be labeled *Sell-out Crowd* and *Smaller Crowd*, and the columns should be labeled *Strong opponent* and *Weak opponent*.

8. Compute and **record** the conditional column proportions for your table.

## Crowd size and Opponent records

9. Does there appear to be an association between crowd size and opponent records? How are they associated? Does it make sense that these variables would be associated?

## Win/Loss and Opponent records

When the Thunder played a strong opponent, they won only 4 of 22 games. When they played a weak opponent, the Thunder won 11 of 19 games.

10. **Record** this data in a two-way table. Now the columns should be labeled *Strong opponent* and *Weak opponent*, and the rows should be labeled *Wins* and *Losses*.

11. Compute and **record** the conditional column proportions for your table.

12. Does there appear to be an association between wins/losses and opponent records?

## Explain and summarize

13. Fill in the blanks: When the Thunder played a strong opponent (instead of a weak opponent), they were (more/less) _____ likely to win, and they were (more/less) _____ to have a large crowd. Therefore __________ is a confounding variable that explains the association between __________ and __________. 


