---
title: "Section 5.1"
date: "October 9, 2019"
output: 
  word_document:
  revealjs::revealjs_presentation:
    fig_width: 14
    fig_height: 7
    self_contained: true
    theme: black
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 600
---

    
# 5.1: Comparing two groups: Categorical Response

## Two Groups: Overview

- In Chapters 1-3, we learned the basic process of statistical inference using tests and confidence intervals.  We did all this by focusing on a single proportion or a single mean.

- In Chapters 5-7, we will take these ideas and extend them to comparing two groups.  We will compare two proportions, two independent means, and paired data.

## Are you having a good day? {data-background="https://media1.giphy.com/media/l396WMbWoqaNeLNPq/source.gif" data-background-size="contain"}

## Are you having a bad day? {data-background="https://media1.giphy.com/media/l396WMbWoqaNeLNPq/source.gif" data-background-size="contain"}

## Does the question matter?

Is the proportion of people who say "no," they are not having a good day the same as the proportion of people who say "yes," they are having a bad day?

## Positive and Negative Perceptions

> - Researchers questioned 30 students (randomly giving them one of the two questions).
> - They then recorded if a positive (yes-good or not-bad) or negative (not-good or yes-bad) response was given.
> - Is this an *observational study* or a *randomized experiment*?
> - Variables? Explanatory/Response?

## Data Frame

```{r, echo=FALSE}
toq <- c(1,1,0,1,1,0,1,1,1,0,1,0,1,0,1,1,0,1,1,1,0,1,0,1,0,1,0,1,0,0)
rsp <- c(1,0,1,1,0,1,1,1,1,0,0,0,1,0,1,1,1,1,1,1,0,1,0,1,0,1,0,1,1,0)
perceptions <- data.frame(Question_type = ifelse(toq==1, "Good Day", "Bad Day"), 
                          Response = ifelse(rsp==1, "Positive", "Negative"))
perceptions$Question_type <- factor(perceptions$Question_type, levels = c("Good Day", "Bad Day"))
perceptions$Response <- factor(perceptions$Response, levels=c("Positive", "Negative"))
perceptions
```

## Two-Way Table

```{r}
ptable <- t(table(perceptions))
ptable
prop.table(ptable,2)
```

## Segmented Bar Graphs

```{r, echo=FALSE, fig.width=5, fig.height=5, fig.align="center"}
barplot(prop.table(ptable,2), legend = rownames(ptable), width=c(0.5,0.5),xlim=c(0,3), space=0.8,col=c("green", "purple"))
```

## Statistic

```{r, echo=FALSE}
ptable <- t(table(perceptions))
ptable
prop.table(ptable,2)
```

**Difference of proportions** of positive responses:
\[
\hat{p}_1 - \hat{p}_2 \approx 0.83 - 0.33 = 0.5
\]

## Another Statistic

```{r, echo=FALSE}
ptable <- t(table(perceptions))
ptable
prop.table(ptable,2)
```

**Relative Risk** 
\[
\frac{\hat{p}_1}{\hat{p}_2} \approx \frac{0.83}{0.33} = 2.5
\]

# Exploration 5.1

## The case of Nurse Kristen Gilbert {data-background="https://www.southcoasttoday.com/storyimage/NB/20010315/NEWS/303159965/AR/0/AR-303159965.jpg" data-background-size="contain"}

## Preview

Was nurse Gilbert killing patients?

1. Observational units: the 1,641 8-hour shifts.
2. Two categorical variables: Gilbert on shift (yes/no), Patient died (yes/no)
3. Explanatory: Gilbert on shift. Response: Patient died.
4. Observational study.

## 

The researchers found that Gilbert worked on 257 of the 1,641 shifts, and she did not work on 1,384 of them. They also found that a patient died during 74 of the shifts, and no death occurred on the other 1,567 shifts. Among the 74 shifts during which a patient died, 40 were shifts on which Gilbert worked and 34 were shifts on which Gilbert did not work.

1. **Record** a two-way table from this data.

|             | Gilbert worked | Gilbert did not work | Total |
| ----------- | -------------- | -------------------- | ----- |
| Patient died |  |  |  |
| No patients died |  |  |  |
| Total |  |  |  |

## Conditional proportions?

2. We want to know the conditional proportion of Gilbert shifts on which a patient died. Do we want column proportions or row proportions? **Record** a second table of these proportions.

## Segmented bar graph

3. **Presenter:** On the board, draw and label a segmented bar graph. The bars should be the explanatory variable, and the segments should be the proportions of the response variable.

4. Does there appear to be a tendency for Gilbert shifts to have at least one death more often than non-Gilbert shifts? Explain.

5. Without doing any further analysis, do you consider the difference between the conditional proportions to be striking, worth reporting to a jury?

## Statistics

6. **Record** the difference of proportions $\hat{p}_1 - \hat{p}_2$.

7. **Record** the relative risk $\frac{\hat{p}_1}{\hat{p}_2}$. **Record** a sentence interpreting this ratio value, as the prosecution might present to a jury. Does the value of this ratio strike you as noteworthy?

8. Suggest an alternative explanation that the defense attorneys might use to explain the large difference in death proportions between the two groups.

## Data Set #2

9. Suppose that the data had turned out as follows:

|             | Gilbert worked | Gilbert did not work | Total |
| ----------- | -------------- | -------------------- | ----- |
| Patient died | 100 | 357 | 457 |
| No patients died | 157 | 1027 | 1184 |
| Total | 257 | 1384 | 1641 |

**Record** $\hat{p}_1 - \hat{p}_2$ and  $\frac{\hat{p}_1}{\hat{p}_2}$ for this data, and compare the values you got for the original data. Which data set (#1 or #2) do you think would be more incriminating? Why? **Operator:** Use [this form](https://goo.gl/forms/7PFMg6T7nJhXP49y1) to share your group's conclusion in one or two sentences.

# Simulation-based comparison of two proportions

## Hypotheses

* Null: Patient deaths are not associated to whether Gilbert was working.
* Alternative: On Gilbert's shifts, patients more likely to die.

\[
H_0: \pi_1 - \pi_2 = 0 \\
H_a: \pi_1 - \pi_2 > 0 
\]

## Chance model

- Could the different death rates be due to chance?
- If deaths just occur randomly, how likely is the data that we observed?

## Chance model $\rightarrow$ null distribution

>- Assume the death rate proportions are the same: $\pi_1 = \pi_2 = 74/1641 \approx 0.045$, whether or not Gilbert was working. So there are <span style="color:cyan">74 "death shifts"</span> and <span style="color:lightgreen">1567 "no-death shifts,"</span> for a total of 1641 shifts.
>- Randomly assign 257 of these shifts to Gilbert. (2-way table)
>- Compute the simulated $\hat{p}_1-\hat{p}_2$. This gives us one point on the dotplot of the null distribution. 
>- See how often something as extreme as our observed $\hat{p}_1-\hat{p}_2 = 0.131$ happens. 

## Two Proportion applet

Let's enter the (real) Gilbert data into the [Two Proportion Applet](http://www.rossmanchance.com/applets/ChiSqShuffle.html?hideExtras=2).

