---
title: "Section 5.3"
date: "October 14, 2019"
output: 
  word_document:
  revealjs::revealjs_presentation:
    fig_width: 14
    fig_height: 7
    self_contained: true
    theme: black
    highlight: zenburn
    css: slidesdjh.css
    center: false
    transition: slide
    reveal_options:
      controls: true
      progress: false
      width: 1080
      height: 600
---

# Pairs today

## Please sit in pairs

*One person in each pair needs a screen* (phone/tablet/laptop).

Sign in on the log sheet. Choose a role:

- The **Recorder** records the indicated answers on the log sheet.
- The **Operator** is in charge of doing calculations using the applets, and other computer tasks. The Operator needs a screen (either a phone or a laptop/tablet).

# Simulation vs. Theory

## Overview: Simulation vs. Theory

- Simulation always works.
- Under certain conditions, we can predict the result of simulation using theory.

# 5.3: Comparing two Proportions: Theory-based approach

## Smoking During Pregnancy

<iframe width="560" height="315" src="https://www.youtube-nocookie.com/embed/C_hg7-7grZU?rel=0&amp;end=57" frameborder="0" allow="autoplay; encrypted-media" allowfullscreen></iframe>

## Smoking and Babies Gender

How does parents’ behavior affect the sex of their children? Fukuda et al., 2002 (Japan) found the following: 

- 255 of 565 births (45.1%) where both parents smoked more than a pack a day were boys. 
- 1975 of 3602 births (54.8%) where both parents did not smoke were boys.

Other studies have shown a reduced male to female birth ratio where high concentrations of other environmental chemicals are present (e.g. industrial pollution, pesticides).

## Smoking/Gender data

```{r, echo=FALSE, fig.height=4, fig.width=5}
smoking <- as.table(matrix(c(1975, 255, 1627, 310), byrow=TRUE, nrow=2))
rownames(smoking) <- c("Boy", "Girl")
colnames(smoking) <- c("Nonsmokers", "Smokers")
addmargins(smoking)
prop.table(smoking, 2)
```

## Segmented Bar Graph

```{r, echo=FALSE, fig.height=4, fig.width=5, fig.align='center'}
barplot(prop.table(smoking,2), legend = rownames(smoking), width=c(0.5,0.5),xlim=c(0,3), space=1.1,col=c("lightblue", "pink"))
```



## Simulation-based approach

```{r, echo=FALSE}
smoking
prop.table(smoking, 2)
```

\[ 
\mbox{test statistic} = \hat{p}_1 - \hat{p}_2 \approx 0.5483 - 0.4513 \approx 0.097
\]

\[
H_0: \pi_1 - \pi_2 = 0 \\
H_a: \pi_1 - \pi_2 \neq 0
\]

Let's enter the data into the [Two Proportion Applet](http://www.rossmanchance.com/applets/ChiSqShuffle.html).

## Theory-based approach

- The null distribution is approximately normal. 
- **Validity condition:** At least 10 observations in each cell.
- Need standard error of difference of proportions. Assuming $\pi_1 = \pi_2$, 

\[
\mbox{standard error} \approx \sqrt{\hat{p}(1-\hat{p})\left(\frac{1}{n_1} + \frac{1}{n_2} \right)} 
\]

where $n_1$ and $n_2$ are the sizes of the two groups, and $\hat{p}$ is the pooled proportion of "successes" in both groups.

## Theory-based $z$-Score

When the sample sizes are large enough, the standardized statistic is

\[
\begin{align}
z &= \frac{\mbox{observed statistic} - \mbox{null value}}{\mbox{standard error of statistic}}  \\ &\approx \frac{(\hat{p}_1 - \hat{p}_2) - 0}{\sqrt{\hat{p}(1-\hat{p})\left(\frac{1}{n_1} + \frac{1}{n_2} \right)}}
\end{align}
\]

## Theory-based $z$-Score

For our data:

\[
\begin{align}
z &= \frac{(\hat{p}_1 - \hat{p}_2) - 0}{\sqrt{\hat{p}(1-\hat{p})\left(\frac{1}{n_1} + \frac{1}{n_2} \right)}} \\
&\approx \frac{(0.548 - 0.451) - 0}{\sqrt{0.535(1-0.535)\left(\frac{1}{3602} + \frac{1}{565} \right)}} \\
&\approx 4.30
\end{align}
\]

## Theory-based Applet

Luckily, there's an applet: [Theory-Based Inference](http://www.rossmanchance.com/applets/TBIA.html)

```{r, echo=FALSE}
addmargins(smoking)
```

## Confidence intervals

When computing a confidence interval, there is no $H_0$, so we can't assume that $\pi_1 = \pi_2$. So the standard error is:

\[
\mbox{standard error} \approx \sqrt{\frac{\hat{p}_1(1-\hat{p}_1)}{n_1} + \frac{\hat{p}_2(1-\hat{p}_2)}{n_2}} 
\]

So a confidence interval for $\pi_1 - \pi_2$ has the form

\[
(\hat{p}_1-\hat{p}_2) \pm
\mbox{multiplier} \times \sqrt{\frac{\hat{p}_1(1-\hat{p}_1)}{n_1} + \frac{\hat{p}_2(1-\hat{p}_2)}{n_2}} 
\]

where the multiplier depends on the confidence level (e.g., 1.96 for 95% confidence).

## Use software to get confidence intervals

For example, use the [Theory-Based Inference](http://www.rossmanchance.com/applets/TBIA.html) applet.

## Using R for theory-based approach

```{r}
prop.test(x = c(1975, 255), n = c(3602, 565))
```


## What you should know

- Identify when a theory-based approach would be **valid** to find the p-value or a confidence interval when evaluating the relationship between two categorical variables.
- Use the Theory-Based Inference applet or R to find theory based p-values and confidence intervals.
- Understand the impacts of confidence level and sample size on confidence interval width the difference in two proportions.
- Understand the impacts of the difference in the sample proportions and sample size on the p-value. 

# Exploration 5.3

## **Blood Donation** {data-background="https://news.psu.edu/sites/default/files/styles/threshold-992/public/BloodDonorNR.jpg"}

<span style="color:darkblue">We are going to look at data from the General Social Survey, which is a **national survey** conducted every two years on a nationwide **random sample** of adult Americans.</span>

## Preview Assignment

Is there a significant difference in blood donation numbers in the US?

<div class="col2">
```{r, echo=FALSE, fig.height=4, fig.width=5}
blood <- as.table(matrix(c(210, 230, 1152, 1106), byrow=TRUE, nrow=2))
rownames(blood) <- c("Donated blood", "Did not donate")
colnames(blood) <- c("Y2002", "Y2004")
addmargins(blood)
print(prop.table(blood, 2), digits=3)
#barplot(prop.table(blood,2), legend = rownames(blood), width=c(0.5,0.5),xlim=c(0,3), space=1.1,col=c("red", "white"))
```

\[
H_0: \pi_{2002} - \pi_{2004} = 0 \\
H_a: \pi_{2002} - \pi_{2004} \neq 0 
\]

\[ 
\begin{align}
\hat{p}_{2002} - \hat{p}_{2004} &\approx -0.018 \\
\mbox{p-value} &\approx 0.21 \\
\mbox{mean}_\mbox{null} &\approx 0.0 \\
\mbox{SD}_\mbox{null} &\approx 0.014
\end{align}
\]
</div>

## Theory-based approach

1. Use the results of the Preview Assignment to **record** a 95% 2SD confidence interval for $\pi_{2002} - \pi_{2004}$. Is zero a plausible value? What does this fact mean in the context of blood donation?
2. Now use the [Theory-Based Inference](http://www.rossmanchance.com/applets/TBIA.html) applet to **record** a 95% confidence interval. How close is your interval to the 2SD interval?
3. **Record** the theory-based p-value, and compare it to what you got in the Preview Assignment.
4. Are the validity conditions met for the theory-based approach? **Record** a sentence explaining why or why not.

## Men vs. Women

Consider the combined data for 2002 and 2004, classified by gender:

```{r, echo=FALSE}
blood2 <- as.table(matrix(c(239, 201, 1032, 1226), byrow=TRUE, nrow=2))
rownames(blood2) <- c("Donated blood", "Did not donate")
colnames(blood2) <- c("Male", "Female")
addmargins(blood2)
```

5. *Analyze these data to address the question of whether American men and women differ with regard to donating blood.* **Record** hypotheses, a p-value, and a confidence interval using theory-based methods, if valid. If you decide that men and women differ significantly, be sure to *estimate by how much* they differ. **Operator:** [Give a summary of your findings](https://goo.gl/forms/YqvQImKqJ3P0mm4n1) in paragraph form.
