---
geometry: left=0.5in, right=0.5in, top=0.8in, bottom=0.4in
classoption: twoside
output: 
  pdf_document:
    keep_tex: true
    includes:
      in_header: latexheader.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
```

# Introduction to Statistics (MA-005) Westmont College, Fall 2022

## Why learn statistics?

In your future vocation, you will be better equipped to be a faithful presence in our modern culture if you know how to analyze data and use it to make decisions. This course will introduce you to the practice of statistics in a wide variety of contexts.  You will learn the fundamental techniques for making inferences from data.  By the end of the course, you should be able to describe data with graphs and numbers, produce data and simulate chance models using randomization, estimate parameters with confidence intervals, assess evidence for a claim with significance tests, and explore correlations using regression.

## What topics will we cover?

We will cover Chapters 1--27 of *Introduction to Modern Statistics* by Mine Çetinkaya-Rundel and Johanna Hardin. This book is freely available at [`https://openintro-ims.netlify.app`](https://openintro-ims.netlify.app/). Here's an overview:

\begin{description}[noitemsep]
\item[Introduction to data.] Data structures, variables, summaries, graphics, and basic data collection and study design techniques.
\item[Exploratory data analysis.] Data visualization and summarization, with particular emphasis on multivariable relationships.
\item[Regression modeling.] Modeling numerical and categorical outcomes with linear and logistic regression and using model results to describe relationships and make predictions.
\item[Foundations for inference.] Case studies are used to introduce the ideas of statistical inference with randomization tests, bootstrap intervals, and mathematical models.
\item[Statistical inference.] Further details of statistical inference using randomization tests, bootstrap intervals, and mathematical models for numerical and categorical data, including $t$-tests, $\chi^2$-tests, and ANOVA.
\item[Inferential modeling.] Extending inference techniques presented thus-far to linear and logistic regression settings and evaluating model performance.
\end{description}

Throughout the course we will use the R programming language to manipulate data, produce graphics, and perform computations.

## How are grades determined?

Our typical class meeting will consist of several short mini-lectures and student **participation** in group discussions. On the night before each class meeting there will be a **daily assignment** due on Canvas. Every three weeks there will be an in-class **exam**, and there will be a cumulative **final exam** on the scheduled date. The following table shows how these assessments are weighted to determine your final grade.

```{r echo=FALSE}
kbl(data.frame(x = c("Exams (4)", "Final Exam", "Daily Assignments", "Participation"),
                 y = c("15% each", "20%", "15%", "5%")), col.names = NULL, booktabs = TRUE)
#kbl(data.frame(x = c("Exams (3)", "Final Exam", "R Projects", "Daily Homework", "Participation"),
#                 y = c("15% each", "20%", "15%", "15%", "5%")), col.names = NULL, booktabs = TRUE)
```

Grades are based on a 90/80/70/60 scale, with $+/-$'s within 3 percent of each letter-grade cutoff. Due dates will appear on Canvas, where you can also keep track of your progress. 

## What other policies should students be aware of?

If you miss a significant number of classes, you will almost definitely do poorly in this class.  If you miss more than six classes without a valid excuse, I reserve the right to terminate you from the course with a grade of~F.\hspace{4pt}  Work missed (including tests) without a valid excuse will receive a zero.

I expect you to check your email on a regular basis.  If you use a non-Westmont email account, please forward your Westmont email to your preferred account. I'll send out notices on Canvas, so make sure you receive Canvas notifications in your email.

Learning communities function best when students have academic integrity.  Cheating is primarily an offense against your classmates because it undermines our learning community.  Therefore, dishonesty of any kind may result in loss of credit for the work involved and the filing of a report with the Provost's Office. Major or repeated infractions may result in dismissal from the course with a grade of F. Be familiar with the College's plagiarism policy, found at \url{https://www.westmont.edu/office-provost/academic-program/academic-integrity-policy}.

In particular, providing someone with an electronic copy of your work is a breach of the academic integrity policy. Do not email, post online, or otherwise disseminate any of the work that you do in this class. You may work with others on the assignments, but make sure that you write or type up your own answers yourself. You are on your honor that the work you hand in represents your own understanding.

\clearpage

## Other Information

\begin{description} 

\item[Professor:] David J. Hunter, Ph.D.
  (\verb!dhunter@westmont.edu!). Student hours are in Winter Hall 303 
  from 1:30--4pm on Tuesdays and Thursdays.

\item[Accommodations for Students with Disabilities:] Students who have been diagnosed with a disability (learning, physical or psychological) are strongly encouraged to contact the Disability Services office as early as possible to discuss appropriate accommodations for this course. Formal accommodations will only be granted for students whose disabilities have been verified by the Disability Services office.  These accommodations may be necessary to ensure your equal access to this course.  Please contact Sheri Noble, Director of Disability Services (310A Voskuyl Library, 565-6186, \href{mailto:snoble@westmont.edu}{\tt snoble@westmont.edu}) or visit \url{https://www.westmont.edu/disability-services} for more information.

\item[General Education: ]
This course fulfills the Quantitative and Analytical Reasoning (QAR) requirement because it emphasizes understanding and communication of numeric data including the computation and interpretation of summative statistics and the presentation and interpretation of graphical representations of data. A core focus of the course is the explicit study of quantitative and analytic methods. This course also fulfills Reasoning Abstractly (RA) because it focuses on critical and analytical reasoning about non-empirical, abstract concepts, objects and structures. You will learn to understand and evaluate abstract arguments and explanations, analyze abstract concepts and solve abstract problems.  Students completing this course will be able to:
\begin{itemize}[noitemsep]
    \item interpret numeric data, summative statistics and graphical representations (QAR);
    \item reflect on the strengths and weaknesses of particular quantitative models or methods as tools in the natural and social sciences (QAR);
    \item be able to interpret, reflect on, and use quantitative models and data in public, vocational, and/or private decision making (QAR);
    \item identify instances of abstract reasoning about abstract objects or concepts (in the form of arguments, explanations, proofs, analyses, modeling, or processes of problem solving) (RA);
    \item construct an instance of valid reasoning about abstract objects or concepts (in the form of arguments, explanations, proofs, analyses, modeling, or processes of problem solving) (RA);
    \item distinguish valid forms of reasoning about abstract objects or concepts (in the form of arguments, explanations, proofs, analyses, modeling, or processes of problem solving) from invalid and/or fallacious
forms of reasoning (RA).
\end{itemize}

\item[Program and Institutional Learning Outcomes:] The
         mathematics department at Westmont College has formulated the
         following learning outcomes for all of its classes. (PLO's)
\begin{enumerate}[noitemsep]
\item Core Knowledge: Students will demonstrate knowledge of the
                  main concepts, skills, and facts of the discipline of
                  mathematics.
\item Communication: Students will be able to communicate mathematical ideas
     following the standard conventions of writing or speaking in the
     discipline.
\item Creativity: Students will demonstrate the ability to formulate and make
     progress toward solving non-routine problems.
\item Christian Connection: Students will incorporate their mathematical skills
     and knowledge into their thinking about their vocations as followers of
     Christ.
         \end{enumerate}[noitemsep]
         In addition, the faculty of Westmont College have established common
         learning outcomes for all courses at the institution
         (ILO's). These outcomes are summarized as follows:
(1) Christian Understanding, Practices, and Affections,
(2) Global Awareness and Diversity,
(3) Critical Thinking,
(4) Quantitative Literacy,
(5) Written Communication,
(6) Oral Communication, and
(7) Information Literacy.

\item[Course Learning Outcomes:] The above outcomes are reflected in the
     particular learning outcomes for this course.
     After taking this course, you should be able
     to:
    \begin{itemize}[noitemsep]
        \item Demonstrate mastery of fundamental concepts of statistics. (PLO 1, ILOs 3,4)
        \item Describe mathematical models and structures according to the
             standards of the discipline. (PLO 2,
              ILOs 3,5)
        \item Present mathematical constructions, computations, and arguments orally, with
              clarity and accuracy. (PLO 2, ILO~6)
        \item Construct solutions to novel mathematical problems,
               demonstrating perseverance in the face of open-ended or
               partially-defined contexts. (PLO 3, ILO 3)
        \item Explain the connection between your personal mathematical
             development and your professional calling. (PLO 4, ILO
             1)
    \end{itemize}
These outcomes will be assessed by group activities, written assignments, and exams, as described above.

\end{description}
